---
title: Reshaping Data with tidyr
author: ''
date: '2021-10-04'
slug: reshaping-data-with-tidyr
categories: []
tags:
  - data_wrangling
  - dataset_preparation
---

```{r include=FALSE}
library(tidyverse)

netflix_df <- read_csv("netflix_df.csv")
phone_nr_df <- read_csv("phone_nr_df.csv")
tvshow_df <- read_csv("tvshow_df.csv")
drink_df <-  read_csv("drink_df.csv")
director_df <-  read_csv("director_df.csv")
sales_df <-  read_csv("sales_df.csv")
country_to_continent_df <-  read_csv("country_to_continent_df.csv")
nuke_df <-  read_csv("nuke_df.csv")
obesity_df <-  read_csv("obesity_df.csv")
nuke_df2 <-  read_csv("nuke_df2.csv")
bond_df <-  read_csv("bond_df.csv")
olympic <- readRDS("olympic_medals_top_10.rds")
```


Data in the wild can be scary—when confronted with a complicated and messy dataset you may find yourself wondering, where do I even start? The tidyr package allows you to wrangle such beasts into nice and tidy datasets. Inaccessible values stored in column names will be put into rows, JSON files will become data frames, and missing values will never go missing again. You'll practice these techniques on a wide range of messy datasets, learning along the way how many dogs the Soviet Union sent into space and what bird is most popular in New Zealand. With the tidyr package in your tidyverse toolkit, you'll be able to transform almost any dataset in a tidy format which will pay-off during the rest of your analysis.

## 1- Tidy data

You'll be introduced to the concept of tidy data which is central to this course. In the first two lessons, you'll jump straight into the action by separating messy character columns into tidy variables and observations ready for analysis. In the final lesson, you'll learn how to overwrite and remove missing values.

## 1-3 Multiple variables per column

Being a busy person, you don't want to spend too much time on Netflix, so you decide to crunch some numbers on TV show and movie durations before deciding what to watch. You've managed to obtain a dataset named netflix_df, but its duration column has an issue. It contains strings with both a value and unit of duration ("min" or "Season").

You'll tidy this dataset so that each variable gets its own column.

```{r}
netflix_df %>% 
  # Split the duration column into value and unit columns
  separate(duration, into = c("value", "unit"), sep = " ", convert = TRUE)
```

## 1-5 International phone numbers

You work for a multinational company that uses auto-dialer software to contact its customers. When new customers subscribe online they are asked for a phone number but they often forget to add the country code needed for international calls. You were asked to fix this issue in the database. You've been given a data frame with national numbers and country codes named phone_nr_df. Now you want to combine the country_code and national_number columns to create valid international numbers.

```{r}
phone_nr_df

phone_nr_df %>%
  # Unite the country_code and national_number columns
  unite("international_number", country_code, national_number, sep = " ")
```

## 1-6 Extracting observations from values

You're given a sample of the Netflix dataset containing TV shows and their casts called tvshow_df. You want to learn which six actors have the most appearances.

However, the dataset only has one row per TV show, and multiple actors are listed in the cast column.

Transform the data so that for each TV show, every actor has a row. The number of appearances will be calculated for you.

```{r}
tvshow_df

tvshow_df %>% 
  # Separate the actors in the cast column over multiple rows
  separate_rows(cast, sep = ", ", convert = TRUE) %>% 
  rename(actor = cast) %>% 
  count(actor, sort = TRUE) %>% 
  head()
```

## 1-7 Separating into columns and rows

Remember the drink ingredients data from the video? You've been given an similar version (drink_df) that also includes quantities and units. Now you want to create an overview of how much of each ingredient you should buy to make these drinks.


```{r}
drink_df


drink_df %>% 
  # Separate the ingredients over rows
  separate_rows(ingredients, sep = "; ") %>% 
  # Separate ingredients into three columns
  separate(
    ingredients, 
    into = c("ingredient", "quantity", "unit"), 
    sep = " ", 
    convert = TRUE
  ) %>% 
  # Group by ingredient and unit
  group_by(ingredient, unit) %>% 
  # Calculate the total quantity of each ingredient
  summarize(quantity = sum(quantity))
```

## 1-9 And the Oscar for best director goes to ... <NA>

You're working on a sample of the Netflix dataset pre-loaded as director_df. This time, the data frame contains just the directors and movie titles. Your goal is to identify the directors who created the most movies. Since the director column contains multiple names, you'll first separate its values over multiple rows and then count the directors.

Since you don't want movies without directors polluting your overview, you'll apply the drop_na() function.

```{r}
director_df %>% 
  # Spread the director column over separate rows
  separate_rows(director, sep = ", ") %>% 
  # Count the number of movies per director
  count(director, sort=TRUE)
```

```{r}
director_df %>% 
  # Drop rows with NA values in the director column
  drop_na() %>% 
  # Spread the director column over separate rows
  separate_rows(director, sep = ", ") %>% 
  # Count the number of movies per director
  count(director, sort = TRUE)
```

## 1-10 Imputing sales data

You've been asked to create a report that allows management to compare sales figures per quarter for two years. The problem is that the dataset (sales_df) contains missing values. You'll need to impute the values in the year column so that you can visualize the data.

```{r}
sales_df %>% 
  # Impute the year column
  fill(year, .direction = "up") %>%
  # Create a line plot with sales per quarter colored by year.
  ggplot(aes(x = quarter, y = sales, color = as.factor(year), group = year)) +
  geom_line()
```

## 1-11 Nuclear bombs per continent

Since WWII, a number of nations have been detonating nuclear bombs for military research. A tally of bombs detonated per nation has been calculated from the Nuclear Explosion DataBase (NEDB) and provided as nuke_df. You are interested in finding out how many bombs have been detonated by nations grouped per continent. To achieve this goal, nuke_df will be joined to country_to_continent_df which is a mapping of nation to continent. You will need to overwrite missing values with zeros so that you can create a nice plot.

The dplyr and ggplot2 packages have been pre-loaded for you.

Side note 1: Bombs detonated by the Soviet Union were attributed to the Russian Federation.

Side note 2: The Russian Federation is solely mapped to Europe for simplicity.

```{r}
country_to_continent_df %>% 
  left_join(nuke_df, by = "country_code") %>%  
  # Impute the missing values in the n_bombs column with 0L
  replace_na(list(n_bombs = 0L)) %>% 
  # Group the dataset by continent
  group_by(continent) %>% 
  # Sum the number of bombs per continent
  summarize(n_bombs_continent = sum(n_bombs)) %>% 
  # Plot the number of bombs per continent
  ggplot(aes(continent, n_bombs_continent)) +
  geom_col()
```

## 2-2 Nuclear bombs per country

You've been given a version of the Nuclear Explosion DataBase (NEDB) where country names are specified in the column headers (nuke_df). You want to visualize how many nukes were detonated per year per country. You'll need to pivot the data and replace NA values first.

```{r}
nuke_df2

nuke_df2 %>% 
  # Pivot the data to a longer format
  pivot_longer(-year)
```

```{r}
nuke_df2 %>% 
  # Pivot the data to a longer format
  pivot_longer(
    -year, 
    # Overwrite the names of the two new columns
    names_to = "country", 
     values_to = "n_bombs"
  ) 
```

```{r}
nuke_df2 %>% 
  # Pivot the data to a longer format
  pivot_longer(
    -year, 
    # Overwrite the names of the two new columns
    names_to = "country", 
    values_to = "n_bombs"
  ) %>%
  # Replace NA values for n_bombs with 0L
   replace_na(list(n_bombs = 0L))
```

```{r}
nuke_df2 %>% 
  # Pivot the data to a longer format
  pivot_longer(
    -year, 
    # Overwrite the names of the two new columns
    names_to = "country", 
    values_to = "n_bombs"
  ) %>% 
  # Replace NA values for n_bombs with 0L
  replace_na(list(n_bombs = 0L)) %>% 
  # Plot the number of bombs per country over time
  ggplot(aes(year,n_bombs, color=country)) +
  geom_line()
```


## 2-3 WHO obesity per country

According to the World Health Organization (WHO), worldwide obesity has nearly tripled since 1975. You're interested in the severity of this global health issue per country and whether males and females are affected differently. You'll use the WHO's obesity data (obesity_df) to investigate this issue. The data holds the percentage of females, males, and both sexes combined that are considered obese (BMI > 30) per country.

You want to create a scatterplot where, per nation, you can see the obesity data colored differently for females and males. This implies that sex should become a variable with its own column.

```{r}
obesity_df

obesity_df %>% 
  # Pivot the male and female columns
  pivot_longer(c(male,female), names_to = "sex", 
    values_to = "pct_obese")
```

```{r}
obesity_df %>% 
  # Pivot the male and female columns
  pivot_longer(c(male, female),
               names_to = "sex",
               values_to = "pct_obese") %>% 
  # Create a scatter plot with pct_obese per country colored by sex
  ggplot(aes(x = pct_obese, color = sex,
             y = forcats::fct_reorder(country, both_sexes))) +
  geom_point() +
  scale_y_discrete(breaks = c("India", "Nauru", "Cuba", "Brazil",
                              "Pakistan", "Gabon", "Italy", "Oman",
                              "China", "United States of America")) +
  labs(x = "% Obese", y = "Country")
```


## 2-4 Bond... James Bond

You've been given a James Bond movie dataset (bond_df) and want to visualize the number of movies that Bond actors have featured in per decade. However, the data is in an untidy format with the decade values captured in the column headers. You'll tidy this dataset to give each variable its own column.

```{r}
bond_df %>% 
  # Pivot the data to long format
  pivot_longer(
    -Bond, 
    # Overwrite the names of the two newly created columns
    names_to = "decade", 
    values_to = "n_movies", 
    # Drop na values
    values_drop_na = TRUE, 
    # Transform the decade column data type to integer
   names_transform= list(decade = as.integer)
  ) %>% 
  ggplot(aes(x = decade + 5, y = n_movies, fill = Bond))+
  geom_col()
```

## 2-6 New-Zealand's bird of the year

Every year New Zealanders vote en masse to decide which species gets the bird of the year trophy. The contest is organized by the Forest & Bird agency which allows each person to give points to up to five birds (first pick gets 5 points, second 4, …). Your job is to decide this year's winner from the messy dataset that's been pre-loaded for you as bird_df.

```{r}

```

## 2-7 Big tech stock prices

You're an analyst at an investment firm and want to visualize the weekly closing prices of five big tech firms' stocks. However, the dataset you've been handed (stock_df) is messy and has the year and week variables stored in the column headers. You'll pivot this data into a tidy format, extract the variables from the headers, and create a line plot.

```{r}

```

## 2-9 Soviet space dogs, the dog perspective

You'll be working on an pre-processed sample of the USSR space dogs database compiled by Duncan Geere and pre-loaded for you as space_dogs_df. Each of the 42 rows in this dataset represents a test rocket launch which had one or two very brave dogs on board.

Your goal is to reshape this dataset so that for each launch, each dog has a row.

The challenge is that in the column headers (name_1, name_2, gender_1, and gender_2), the part before the _ separator can point to two different variables (name and gender), while the second part always points to the dog ID (1st or 2nd dog).

```{r}

```

## 2-10 WHO obesity vs. life expectancy

You've been given a sample of WHO data (who_df) with obesity percentages and life expectancy data per country, year, and sex. You want to visually inspect the correlation between obesity and life expectancy.

However, the data is very messy with four variables hidden in the column names. Each column name is made up of three parts separated by underscores: Values for the year, followed by those for sex, and then values for either pct.obese or life.exp. Since the third part of the column name string holds two variables you'll need to use the special ".value" value in the names_to argument.

You'll pivot the data into a tidy format and create the scatterplot.

```{r}

```

## 2-11 Uncounting observations

You've found the job of your dreams providing technical support for a dog breed beauty contest. The jury members want a spreadsheet with the breed and id of each participating dog so that they can add the scores later on. You've only been given the number of participants per dog breed (dog_df) so you decide to use your tidyr skills to create the desired result.

```{r}

```

## 2-13 Soviet space dogs, the flight perspective

Remember the USSR space dogs dataset1? You changed it to a long format so that for every dog in every rocket launch, there was a row. Suppose you're given this tidy dataset and are asked to answer the question, "In what percentage of flights were both dogs of the same gender?"

You'll reshape and investigate space_dogs_df to find the answer.

```{r}

```

## 2-14 Planet temperature & distance to the Sun

The intensity of light radiated by a light source follows an inverse square relationship with the distance it has traveled.

https://en.wikipedia.org/wiki/Inverse-square_law 

You wonder if you could observe this trend in the temperature of the planets in our Solar System given their distance to the Sun. You'll use the planet_df dataset from the devstronomy project to investigate this.

```{r}

```

## 2-15 Transposing planet data

You're again working on a planet dataset derived from the devstronomy project. This time, you're interested in the correlation between the diameter of a planet and the number of moons circling it.

However, the dataset (planet_df) has a row for each variable and a column for each planet (observation). You'll transpose this data in two steps and then create a plot to inspect the correlation.

```{r}

```

