---
title: Reshaping Data with tidyr
author: ''
date: '2021-10-04'
slug: reshaping-data-with-tidyr
categories: []
tags:
  - data_wrangling
  - dataset_preparation
---

```{r include=FALSE}
library(tidyverse)

netflix_df <- read_csv("netflix_df.csv")
phone_nr_df <- read_csv("phone_nr_df.csv")
tvshow_df <- read_csv("tvshow_df.csv")
drink_df <-  read_csv("drink_df.csv")
director_df <-  read_csv("director_df.csv")
olympic <- readRDS("olympic_medals_top_10.rds")
```


Data in the wild can be scary—when confronted with a complicated and messy dataset you may find yourself wondering, where do I even start? The tidyr package allows you to wrangle such beasts into nice and tidy datasets. Inaccessible values stored in column names will be put into rows, JSON files will become data frames, and missing values will never go missing again. You'll practice these techniques on a wide range of messy datasets, learning along the way how many dogs the Soviet Union sent into space and what bird is most popular in New Zealand. With the tidyr package in your tidyverse toolkit, you'll be able to transform almost any dataset in a tidy format which will pay-off during the rest of your analysis.

## 1- Tidy data

You'll be introduced to the concept of tidy data which is central to this course. In the first two lessons, you'll jump straight into the action by separating messy character columns into tidy variables and observations ready for analysis. In the final lesson, you'll learn how to overwrite and remove missing values.

## 1-3 Multiple variables per column

Being a busy person, you don't want to spend too much time on Netflix, so you decide to crunch some numbers on TV show and movie durations before deciding what to watch. You've managed to obtain a dataset named netflix_df, but its duration column has an issue. It contains strings with both a value and unit of duration ("min" or "Season").

You'll tidy this dataset so that each variable gets its own column.

```{r}
netflix_df %>% 
  # Split the duration column into value and unit columns
  separate(duration, into = c("value", "unit"), sep = " ", convert = TRUE)
```

## 1-5 International phone numbers

You work for a multinational company that uses auto-dialer software to contact its customers. When new customers subscribe online they are asked for a phone number but they often forget to add the country code needed for international calls. You were asked to fix this issue in the database. You've been given a data frame with national numbers and country codes named phone_nr_df. Now you want to combine the country_code and national_number columns to create valid international numbers.

```{r}
phone_nr_df

phone_nr_df %>%
  # Unite the country_code and national_number columns
  unite("international_number", country_code, national_number, sep = " ")
```

## 1-6 Extracting observations from values

You're given a sample of the Netflix dataset containing TV shows and their casts called tvshow_df. You want to learn which six actors have the most appearances.

However, the dataset only has one row per TV show, and multiple actors are listed in the cast column.

Transform the data so that for each TV show, every actor has a row. The number of appearances will be calculated for you.

```{r}
tvshow_df

tvshow_df %>% 
  # Separate the actors in the cast column over multiple rows
  separate_rows(cast, sep = ", ", convert = TRUE) %>% 
  rename(actor = cast) %>% 
  count(actor, sort = TRUE) %>% 
  head()
```

## 1-7 Separating into columns and rows

Remember the drink ingredients data from the video? You've been given an similar version (drink_df) that also includes quantities and units. Now you want to create an overview of how much of each ingredient you should buy to make these drinks.


```{r}
drink_df


drink_df %>% 
  # Separate the ingredients over rows
  separate_rows(ingredients, sep = "; ") %>% 
  # Separate ingredients into three columns
  separate(
    ingredients, 
    into = c("ingredient", "quantity", "unit"), 
    sep = " ", 
    convert = TRUE
  ) %>% 
  # Group by ingredient and unit
  group_by(ingredient, unit) %>% 
  # Calculate the total quantity of each ingredient
  summarize(quantity = sum(quantity))
```

## 1-9 And the Oscar for best director goes to ... <NA>

You're working on a sample of the Netflix dataset pre-loaded as director_df. This time, the data frame contains just the directors and movie titles. Your goal is to identify the directors who created the most movies. Since the director column contains multiple names, you'll first separate its values over multiple rows and then count the directors.

Since you don't want movies without directors polluting your overview, you'll apply the drop_na() function.

```{r}
director_df %>% 
  # Spread the director column over separate rows
  separate_rows(director, sep = ", ") %>% 
  # Count the number of movies per director
  count(director, sort=TRUE)
```

```{r}
director_df %>% 
  # Drop rows with NA values in the director column
  drop_na() %>% 
  # Spread the director column over separate rows
  separate_rows(director, sep = ", ") %>% 
  # Count the number of movies per director
  count(director, sort = TRUE)
```

## 1-10 Imputing sales data

You've been asked to create a report that allows management to compare sales figures per quarter for two years. The problem is that the dataset (sales_df) contains missing values. You'll need to impute the values in the year column so that you can visualize the data.

```{r}

```

## 1-11 Nuclear bombs per continent

Since WWII, a number of nations have been detonating nuclear bombs for military research. A tally of bombs detonated per nation has been calculated from the Nuclear Explosion DataBase (NEDB) and provided as nuke_df. You are interested in finding out how many bombs have been detonated by nations grouped per continent. To achieve this goal, nuke_df will be joined to country_to_continent_df which is a mapping of nation to continent. You will need to overwrite missing values with zeros so that you can create a nice plot.

The dplyr and ggplot2 packages have been pre-loaded for you.

Side note 1: Bombs detonated by the Soviet Union were attributed to the Russian Federation.

Side note 2: The Russian Federation is solely mapped to Europe for simplicity.

```{r}

```

