---
title: Foundations of Inference
author: ''
date: '2021-10-13'
slug: foundations-of-inference
categories: []
tags:
  - Inference
---

```{r include=FALSE}
library(tidyverse)
library(infer)
```

One of the foundational aspects of statistical analysis is inference, or the process of drawing conclusions about a larger population from a sample of data. Although counter intuitive, the standard practice is to attempt to disprove a research claim that is not of interest. For example, to show that one medical treatment is better than another, we can assume that the two treatments lead to equal survival rates only to then be disproved by the data. Additionally, we introduce the idea of a p-value, or the degree of disagreement between the data and the hypothesis. We also dive into confidence intervals, which measure the magnitude of the effect of interest (e.g. how much better one treatment is than another).

## 1- Introduction to ideas of inference

In this chapter, you will investigate how repeated samples taken from a population can vary. It is the variability in samples that allow you to make claims about the population of interest. It is important to remember that the research claims of interest focus on the population while the information available comes only from the sample data.

## 1-5 Working with the NHANES data

Throughout this chapter, you will use the NHANES dataset from the NHANES R package. The data are collected by the Center for Disease Control (CDC, the national public health institute in the United States) and can be thought of as a random sample of US residents.

Before moving on to investigate particular variables, you'll have an opportunity to briefly explore the data in this exercise.


```{r}
library(NHANES)

# What are the variables in the NHANES dataset?
colnames(NHANES)

# Create bar plot for Home Ownership by Gender
ggplot(NHANES, aes(x = Gender, fill = HomeOwn)) + 
  # Set the position to fill
  geom_bar(position = "fill") +
  ylab("Relative frequencies")


# Density plot of SleepHrsNight colored by SleepTrouble
ggplot(NHANES, aes(x = SleepHrsNight, color = SleepTrouble)) + 
  # Adjust by 2
  geom_density(adjust = 2) + 
  # Facet by HealthGen
  facet_wrap(~ HealthGen)
```

## 1-6 Calculating statistic of interest

Using the NHANES dataset, let's investigate the relationship between gender and home ownership. Remember, more information about the dataset can be found here: 

https://www.rdocumentation.org/packages/NHANES/versions/2.1.0/topics/NHANES

As seen in the video, natural variability can be modeled from shuffling observations around to remove any relationships that might exist in the population. However, before you permute the data, you need to calculate the original observed statistic. In this exercise, you will calculate the difference in proportion of home owners who are men versus women.

Recall that:

%in% returns a logical vector that is TRUE when values on the left hand side are listed on the right hand side.
The mean of a logical vector is the proportion of cases where that vector is TRUE.
fruits <- c("apple", "banana", "cherry")
fruits %in% c("banana", "cherry")
mean(fruits %in% c("banana", "cherry"))

```{r}
# From previous step
homes <- NHANES %>%
# Select Gender and HomeOwn
  select(Gender, HomeOwn) %>%
# Filter for HomeOwn equal to "Own" or "Rent"
  filter(HomeOwn %in% c("Own", "Rent"))

diff_orig <- homes %>%   
  # Group by gender
  group_by(Gender ) %>%
  # Summarize proportion of homeowners
  summarize(prop_own = mean(HomeOwn=="Own")) %>%
  # Summarize difference in proportion of homeowners
  summarize(obs_diff_prop = diff(prop_own)) # male - female
  
# See the result
diff_orig
```

## 1-7 Randomized data under null model of independence

The infer package will allow you to model a particular null hypothesis and then randomize the data to calculate permuted statistics. In this exercise, after specifying your null hypothesis you will permute the home ownership variable 10 times. By doing so, you will ensure that there is no relationship between home ownership and gender, so any difference in home ownership proportion for female versus male will be due only to natural variability.

This exercise will demonstrate the first three steps from the infer package:

specify will specify the response and explanatory variables.
hypothesize will declare the null hypothesis.
generate will generate resamples, permutations, or simulations.
The dplyr, ggplot2, NHANES, and infer packages have been loaded for you and the homes dataset you created in the last exercise is available in your workspace.

```{r}

```


## 1-8 Randomized statistics and dotplot

By permuting the home ownership variable multiple times, you generate differences in proportions that are consistent with the assumption that the variables are unrelated. The statistic of interest is the difference in proportions given by stat = "diff in props". After calculating the randomized statistics, you will plot them in a dotplot.

This exercise shows all four steps from the infer package:

specify will specify the response and explanatory variables.
hypothesize will declare the null hypothesis.
generate will generate resamples, permutations, or simulations.
calculate will calculate summary statistics.
Each step will be covered throughout the course; in this exercise you'll write code for calculate().

The dplyr, ggplot2, NHANES, and infer packages have been loaded for you. Repeat the permuting and plotting with 100 differences in proportions generated by shuffling the HomeOwn variable.

```{r}

```

## 1-9 Randomization density

Using 100 repetitions allows you to understand the mechanism of permuting. However, 100 is not enough to observe the full range of likely values for the null differences in proportions.

Recall the four steps of inference. These are the same four steps that will be used in all inference exercises in this course and future statistical inference courses. Use the names of the functions to help you recall the analysis process.

specify will specify the response and explanatory variables.
hypothesize will declare the null hypothesis.
generate will generate resamples, permutations, or simulations.
calculate will calculate summary statistics.
In this exercise, you'll repeat the process 1000 times to get a sense for the complete distribution of null differences in proportions.

```{r}
# Perform 1000 permutations
homeown_perm <- homes %>%
  # Specify HomeOwn vs. Gender, with `"Own" as success
  specify(HomeOwn ~ Gender, success = "Own") %>%
  # Use a null hypothesis of independence
  hypothesize(null = "independence") %>% 
  # Generate 1000 repetitions (by permutation)
  generate(reps = 1000, type = "permute") %>% 
  # Calculate the difference in proportions (male then female)
  calculate(stat = "diff in props", order = c("male", "female"))

# Density plot of 1000 permuted differences in proportions
ggplot(homeown_perm, aes(x = stat)) + 
  geom_density()
```


## 1-11 Do the data come from the population?

Recall that the observed difference (i.e. the difference in proportions in the homes dataset, shown as the red vertical line) was around -0.0078, which seems to fall below the bulk of the density of shuffled differences. It is important to know, however, whether any of the randomly permuted differences were as extreme as the observed difference.

In this exercise, you'll re-create this dotplot as a density plot and count the number of permuted differences that were to the left of the observed difference.

```{r}

```


## 2- Completing a randomization test: gender discrimination

In this chapter, you will gain the tools and knowledge to complete a full hypothesis test. That is, given a dataset, you will know whether or not is appropriate to reject the null hypothesis in favor of the research claim of interest.

## 2-2- 

```{r}

```

## 3- Hypothesis testing errors: opportunity cost

You will continue learning about hypothesis testing with a new example and the same structure of randomization tests. In this chapter, however, the focus will be on different errors (type I and type II), how they are made, when one is worse than another, and how things like sample size and effect size impact the error rates.

## 3-2- 

```{r}

```

## 4- Confidence intervals

As a complement to hypothesis testing, confidence intervals allow you to estimate a population parameter. Recall that your interest is always in some characteristic of the population, but you only have incomplete information to estimate the parameter using sample data. Here, the parameter is the true proportion of successes in a population. Bootstrapping is used to estimate the variability needed to form the confidence interval.

## 4-2- 

```{r}

```