---
title: Intermediate R
author: ''
date: '2021-10-20'
slug: intermediate-r
categories: []
tags:
  - R
---
```{r include=FALSE}
li_df <- read.csv("li_df.csv")
```

Intermediate R is the next stop on your journey in mastering the R programming language. In this R training, you will learn about conditional statements, loops, and functions to power your own R scripts. Next, make your R code more efficient and readable using the apply functions. Finally, the utilities chapter gets you up to speed with regular expressions in R, data structure manipulations, and times and dates. This course will allow you to take the next step in advancing your overall knowledge and capabilities while programming in R.

## 1- Conditionals and Control Flow

In this chapter, you'll learn about relational operators for comparing R objects, and logical operators like "and" and "or" for combining TRUE and FALSE values. Then, you'll use this knowledge to build conditional statements.

## 1-2 Equality

The most basic form of comparison is equality. Let's briefly recap its syntax. The following statements all evaluate to TRUE (feel free to try them out in the console).

3 == (2 + 1)
"intermediate" != "r"
TRUE != FALSE
"Rchitect" != "rchitect"

Notice from the last expression that R is case sensitive: "R" is not equal to "r". Keep this in mind when solving the exercises in this chapter!

```{r}
# Comparison of logicals
TRUE == FALSE

# Comparison of numerics
-6 * 14 != 17 - 101

# Comparison of character strings
"useR" == "user"

# Compare a logical with a numeric
TRUE == 1 
```


## 1-3 Greater and less than

Apart from equality operators, Filip also introduced the less than and greater than operators: < and >. You can also add an equal sign to express less than or equal to or greater than or equal to, respectively. Have a look at the following R expressions, that all evaluate to FALSE:

(1 + 2) > 4
"dog" < "Cats"
TRUE <= FALSE
Remember that for string comparison, R determines the greater than relationship based on alphabetical order. Also, keep in mind that TRUE is treated as 1 for arithmetic, and FALSE is treated as 0. Therefore, FALSE < TRUE is TRUE.

```{r}
# Comparison of numerics
-6 * 5 + 2 >=  -10 + 1

# Comparison of character strings
"raining" <= "raining dogs"

# Comparison of logicals
TRUE > FALSE
```


## 1-4 Compare vectors

You are already aware that R is very good with vectors. Without having to change anything about the syntax, R's relational operators also work on vectors.

Let's go back to the example that was started in the video. You want to figure out whether your activity on social media platforms have paid off and decide to look at your results for LinkedIn and Facebook. The sample code in the editor initializes the vectors linkedin and facebook. Each of the vectors contains the number of profile views your LinkedIn and Facebook profiles had over the last seven days.

Using relational operators, find a logical answer, i.e. TRUE or FALSE, for the following questions:

On which days did the number of LinkedIn profile views exceed 15?

When was your LinkedIn profile viewed only 5 times or fewer?

When was your LinkedIn profile visited more often than your Facebook profile?


```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Popular days
linkedin > 15

# Quiet days
linkedin <= 5

# LinkedIn more popular than Facebook
facebook < linkedin
```


## 1-5 Compare matrices

R's ability to deal with different data structures for comparisons does not stop at vectors. Matrices and relational operators also work together seamlessly!

Instead of in vectors (as in the previous exercise), the LinkedIn and Facebook data is now stored in a matrix called views. The first row contains the LinkedIn information; the second row the Facebook information. The original vectors facebook and linkedin are still available as well.

When were the views exactly equal to 13? Use the views matrix to return a logical matrix.

For which days were the number of views less than or equal to 14? Again, have R return a logical matrix.

```{r}
# The social data has been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)
views <- matrix(c(linkedin, facebook), nrow = 2, byrow = TRUE)
views
# When does views equal 13?
views == 13

# When is views less than or equal to 14?
views <= 14
```


## 1-7 & and |

Before you work your way through the next exercises, have a look at the following R expressions. All of them will evaluate to TRUE:

TRUE & TRUE
FALSE | TRUE
5 <= 5 & 2 < 3
3 < 4 | 7 < 6
Watch out: 3 < x < 7 to check if x is between 3 and 7 will not work; you'll need 3 < x & x < 7 for that.

In this exercise, you'll be working with the last variable. This variable equals the last value of the linkedin vector that you've worked with previously. The linkedin vector represents the number of LinkedIn views your profile had in the last seven days, remember? Both the variables linkedin and last have been pre-defined for you.

```{r}
# The linkedin and last variable are already defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
last <- tail(linkedin, 1)
last
# Is last under 5 or above 10?
last < 5 | last > 10

# Is last between 15 (exclusive) and 20 (inclusive)?
15 <last & last <= 20
```


## 1-8 & and | (2)
Like relational operators, logical operators work perfectly fine with vectors and matrices.

Both the vectors linkedin and facebook are available again. Also a matrix - views - has been defined; its first and second row correspond to the linkedin and facebook vectors, respectively. Ready for some advanced queries to gain more insights into your social outreach?

```{r}
# The social data (linkedin, facebook, views) has been created for you

# linkedin exceeds 10 but facebook below 10
linkedin > 10 & facebook < 10

# When were one or both visited at least 12 times?
linkedin >= 12 | facebook >= 12

# When is views between 11 (exclusive) and 14 (inclusive)?
views > 11 & views <= 14
```


## 1-10 Blend it all together

With the things you've learned by now, you're able to solve pretty cool problems.

Instead of recording the number of views for your own LinkedIn profile, suppose you conducted a survey inside the company you're working for. You've asked every employee with a LinkedIn profile how many visits their profile has had over the past seven days. You stored the results in a data frame called li_df. This data frame is available in the workspace; type li_df in the console to check it out.

Select the entire second column, named day2, from the li_df data frame as a vector and assign it to second.

Use second to create a logical vector, that contains TRUE if the corresponding number of views is strictly greater than 25 or strictly lower than 5 and FALSE otherwise. Store this logical vector as extremes.
Use sum() on the extremes vector to calculate the number of TRUEs in extremes (i.e. to calculate the number of employees that are either very popular or very low-profile). Simply print this number to the console.

```{r}
# li_df is pre-loaded in your workspace

# Select the second column, named day2, from li_df: second
second <- li_df$day2

# Build a logical vector, TRUE if value in second is extreme: extremes
extremes <- second < 5 | second > 25 

# Count the number of TRUEs in extremes
sum(extremes==TRUE)
```


## 1-12 The if statement

Before diving into some exercises on the if statement, have another look at its syntax:

if (condition) {
  expr
}

Remember your vectors with social profile views? Let's look at it from another angle. The medium variable gives information about the social website; the num_views variable denotes the actual number of views that particular medium had on the last day of your recordings. Both variables have been pre-defined for you.

```{r}
# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Examine the if statement for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
}

# Write the if statement for num_views
if (num_views > 15) {
  print("You are popular!")
}
```


## 1-13 Add an else

You can only use an else statement in combination with an if statement. The else statement does not require a condition; its corresponding code is simply run if all of the preceding conditions in the control structure are FALSE. Here's a recipe for its usage:

if (condition) {
  expr1
} else {
  expr2
}

It's important that the else keyword comes on the same line as the closing bracket of the if part!

Both if statements that you coded in the previous exercises are already available to use. It's now up to you to extend them with the appropriate else statements!

```{r}
# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Control structure for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
} else {
print("Unknown medium")
}


# Control structure for num_views
if (num_views > 15) {
  print("You're popular!")
} else {
 print("Try to be more visible!")
}
```


## 1-14 Customize further: else if

The else if statement allows you to further customize your control structure. You can add as many else if statements as you like. Keep in mind that R ignores the remainder of the control structure once a condition has been found that is TRUE and the corresponding expressions have been executed. Here's an overview of the syntax to freshen your memory:

if (condition1) {
  expr1
} else if (condition2) {
  expr2
} else if (condition3) {
  expr3
} else {
  expr4
}
Again, It's important that the else if keywords comes on the same line as the closing bracket of the previous part of the control construct!

```{r}
# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Control structure for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
} else if (medium == "Facebook") {
  # Add code to print correct string when condition is TRUE
print("Showing Facebook information")
} else {
  print("Unknown medium")
}

# Control structure for num_views
if (num_views > 15) {
  print("You're popular!")
} else if (num_views <= 15 & num_views > 10) {
  # Add code to print correct string when condition is TRUE
print("Your number of views is average")
} else {
  print("Try to be more visible!")
}
```


## 1-15 Take control!

In this exercise, you will combine everything that you've learned so far: relational operators, logical operators and control constructs. You'll need it all!

We've pre-defined two values for you: li and fb, denoting the number of profile views your LinkedIn and Facebook profile had on the last day of recordings. Go through the instructions to create R code that generates a 'social media score', sms, based on the values of li and fb.

```{r}
# Variables related to your last day of recordings
li <- 15
fb <- 9

# Code the control-flow construct
if (li >= 15 & fb >= 15) {
  sms <- 2 * (li + fb)
} else if (li < 10 & fb < 10) {
  sms <- 0.5 * (li + fb)
} else {
  sms <- li + fb
}

# Print the resulting sms to the console
sms
```



## 2- Loops

Loops can come in handy on numerous occasions. While loops are like repeated if statements, the for loop is designed to iterate over all elements in a sequence. Learn about them in this chapter.

## 2-2 

```{r}

```

## 2-2 

```{r}

```

## 2-2 

```{r}

```

## 2-2 

```{r}

```

## 2-2 

```{r}

```

## 2-2 

```{r}

```

## 2-2 

```{r}

```

## 2-2 

```{r}

```

## 2-2 

```{r}

```

## 2-2 

```{r}

```


## 3- Functions

Functions are an extremely important concept in almost every programming language, and R is no different. Learn what functions are and how to use them—then take charge by writing your own functions.

## 3-2 

```{r}

```

## 3-2 

```{r}

```

## 3-2 

```{r}

```

## 3-2 

```{r}

```

## 3-2 

```{r}

```

## 3-2 

```{r}

```

## 3-2 

```{r}

```

## 3-2 

```{r}

```

## 3-2 

```{r}

```

## 3-2 

```{r}

```

## 3-2 

```{r}

```


## 4- The apply family

Whenever you're using a for loop, you may want to revise your code to see whether you can use the lapply function instead. Learn all about this intuitive way of applying a function over a list or a vector, and how to use its variants, sapply and vapply.

## 4-2 

```{r}

```

## 4-2 

```{r}

```


## 4-2 

```{r}

```


## 4-2 

```{r}

```


## 4-2 

```{r}

```


## 4-2 

```{r}

```


## 4-2 

```{r}

```


## 4-2 

```{r}

```


## 4-2 

```{r}

```


## 4-2 

```{r}

```


## 4-2 

```{r}

```


## 4-2 

```{r}

```



## 5- Utilities

Mastering R programming is not only about understanding its programming concepts. Having a solid understanding of a wide range of R functions is also important. This chapter introduces you to many useful functions for data structure manipulation, regular expressions, and working with times and dates.

## 5-2 

```{r}

```

## 5-3 

```{r}

```

## 5-2 

```{r}

```

## 5-2 

```{r}

```

## 5-2 

```{r}

```

## 5-2 

```{r}

```

## 5-2 

```{r}

```

## 5-2 

```{r}

```

## 5-2 

```{r}

```

## 5-2 

```{r}

```

## 5-2 

```{r}

```

## 5-2 

```{r}

```

## 5-2 

```{r}

```

## 5-2 

```{r}

```

