---
title: Intermediate R
author: ''
date: '2021-10-20'
slug: intermediate-r
categories: []
tags:
  - R
---
```{r include=FALSE}
li_df <- read.csv("li_df.csv")

ttt <- matrix(c("0", NA, "X", NA, "0", NA, "X", "0", "X"), nrow=3)
```

Intermediate R is the next stop on your journey in mastering the R programming language. In this R training, you will learn about conditional statements, loops, and functions to power your own R scripts. Next, make your R code more efficient and readable using the apply functions. Finally, the utilities chapter gets you up to speed with regular expressions in R, data structure manipulations, and times and dates. This course will allow you to take the next step in advancing your overall knowledge and capabilities while programming in R.

## 1- Conditionals and Control Flow

In this chapter, you'll learn about relational operators for comparing R objects, and logical operators like "and" and "or" for combining TRUE and FALSE values. Then, you'll use this knowledge to build conditional statements.

## 1-2 Equality

The most basic form of comparison is equality. Let's briefly recap its syntax. The following statements all evaluate to TRUE (feel free to try them out in the console).

3 == (2 + 1)
"intermediate" != "r"
TRUE != FALSE
"Rchitect" != "rchitect"

Notice from the last expression that R is case sensitive: "R" is not equal to "r". Keep this in mind when solving the exercises in this chapter!

```{r}
# Comparison of logicals
TRUE == FALSE

# Comparison of numerics
-6 * 14 != 17 - 101

# Comparison of character strings
"useR" == "user"

# Compare a logical with a numeric
TRUE == 1 
```


## 1-3 Greater and less than

Apart from equality operators, Filip also introduced the less than and greater than operators: < and >. You can also add an equal sign to express less than or equal to or greater than or equal to, respectively. Have a look at the following R expressions, that all evaluate to FALSE:

(1 + 2) > 4
"dog" < "Cats"
TRUE <= FALSE
Remember that for string comparison, R determines the greater than relationship based on alphabetical order. Also, keep in mind that TRUE is treated as 1 for arithmetic, and FALSE is treated as 0. Therefore, FALSE < TRUE is TRUE.

```{r}
# Comparison of numerics
-6 * 5 + 2 >=  -10 + 1

# Comparison of character strings
"raining" <= "raining dogs"

# Comparison of logicals
TRUE > FALSE
```


## 1-4 Compare vectors

You are already aware that R is very good with vectors. Without having to change anything about the syntax, R's relational operators also work on vectors.

Let's go back to the example that was started in the video. You want to figure out whether your activity on social media platforms have paid off and decide to look at your results for LinkedIn and Facebook. The sample code in the editor initializes the vectors linkedin and facebook. Each of the vectors contains the number of profile views your LinkedIn and Facebook profiles had over the last seven days.

Using relational operators, find a logical answer, i.e. TRUE or FALSE, for the following questions:

On which days did the number of LinkedIn profile views exceed 15?

When was your LinkedIn profile viewed only 5 times or fewer?

When was your LinkedIn profile visited more often than your Facebook profile?


```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Popular days
linkedin > 15

# Quiet days
linkedin <= 5

# LinkedIn more popular than Facebook
facebook < linkedin
```


## 1-5 Compare matrices

R's ability to deal with different data structures for comparisons does not stop at vectors. Matrices and relational operators also work together seamlessly!

Instead of in vectors (as in the previous exercise), the LinkedIn and Facebook data is now stored in a matrix called views. The first row contains the LinkedIn information; the second row the Facebook information. The original vectors facebook and linkedin are still available as well.

When were the views exactly equal to 13? Use the views matrix to return a logical matrix.

For which days were the number of views less than or equal to 14? Again, have R return a logical matrix.

```{r}
# The social data has been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)
views <- matrix(c(linkedin, facebook), nrow = 2, byrow = TRUE)
views
# When does views equal 13?
views == 13

# When is views less than or equal to 14?
views <= 14
```


## 1-7 & and |

Before you work your way through the next exercises, have a look at the following R expressions. All of them will evaluate to TRUE:

TRUE & TRUE
FALSE | TRUE
5 <= 5 & 2 < 3
3 < 4 | 7 < 6
Watch out: 3 < x < 7 to check if x is between 3 and 7 will not work; you'll need 3 < x & x < 7 for that.

In this exercise, you'll be working with the last variable. This variable equals the last value of the linkedin vector that you've worked with previously. The linkedin vector represents the number of LinkedIn views your profile had in the last seven days, remember? Both the variables linkedin and last have been pre-defined for you.

```{r}
# The linkedin and last variable are already defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
last <- tail(linkedin, 1)
last
# Is last under 5 or above 10?
last < 5 | last > 10

# Is last between 15 (exclusive) and 20 (inclusive)?
15 <last & last <= 20
```


## 1-8 & and | (2)
Like relational operators, logical operators work perfectly fine with vectors and matrices.

Both the vectors linkedin and facebook are available again. Also a matrix - views - has been defined; its first and second row correspond to the linkedin and facebook vectors, respectively. Ready for some advanced queries to gain more insights into your social outreach?

```{r}
# The social data (linkedin, facebook, views) has been created for you

# linkedin exceeds 10 but facebook below 10
linkedin > 10 & facebook < 10

# When were one or both visited at least 12 times?
linkedin >= 12 | facebook >= 12

# When is views between 11 (exclusive) and 14 (inclusive)?
views > 11 & views <= 14
```


## 1-10 Blend it all together

With the things you've learned by now, you're able to solve pretty cool problems.

Instead of recording the number of views for your own LinkedIn profile, suppose you conducted a survey inside the company you're working for. You've asked every employee with a LinkedIn profile how many visits their profile has had over the past seven days. You stored the results in a data frame called li_df. This data frame is available in the workspace; type li_df in the console to check it out.

Select the entire second column, named day2, from the li_df data frame as a vector and assign it to second.

Use second to create a logical vector, that contains TRUE if the corresponding number of views is strictly greater than 25 or strictly lower than 5 and FALSE otherwise. Store this logical vector as extremes.
Use sum() on the extremes vector to calculate the number of TRUEs in extremes (i.e. to calculate the number of employees that are either very popular or very low-profile). Simply print this number to the console.

```{r}
# li_df is pre-loaded in your workspace

# Select the second column, named day2, from li_df: second
second <- li_df$day2

# Build a logical vector, TRUE if value in second is extreme: extremes
extremes <- second < 5 | second > 25 

# Count the number of TRUEs in extremes
sum(extremes==TRUE)
```


## 1-12 The if statement

Before diving into some exercises on the if statement, have another look at its syntax:

if (condition) {
  expr
}

Remember your vectors with social profile views? Let's look at it from another angle. The medium variable gives information about the social website; the num_views variable denotes the actual number of views that particular medium had on the last day of your recordings. Both variables have been pre-defined for you.

```{r}
# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Examine the if statement for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
}

# Write the if statement for num_views
if (num_views > 15) {
  print("You are popular!")
}
```


## 1-13 Add an else

You can only use an else statement in combination with an if statement. The else statement does not require a condition; its corresponding code is simply run if all of the preceding conditions in the control structure are FALSE. Here's a recipe for its usage:

if (condition) {
  expr1
} else {
  expr2
}

It's important that the else keyword comes on the same line as the closing bracket of the if part!

Both if statements that you coded in the previous exercises are already available to use. It's now up to you to extend them with the appropriate else statements!

```{r}
# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Control structure for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
} else {
print("Unknown medium")
}


# Control structure for num_views
if (num_views > 15) {
  print("You're popular!")
} else {
 print("Try to be more visible!")
}
```


## 1-14 Customize further: else if

The else if statement allows you to further customize your control structure. You can add as many else if statements as you like. Keep in mind that R ignores the remainder of the control structure once a condition has been found that is TRUE and the corresponding expressions have been executed. Here's an overview of the syntax to freshen your memory:

if (condition1) {
  expr1
} else if (condition2) {
  expr2
} else if (condition3) {
  expr3
} else {
  expr4
}
Again, It's important that the else if keywords comes on the same line as the closing bracket of the previous part of the control construct!

```{r}
# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Control structure for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
} else if (medium == "Facebook") {
  # Add code to print correct string when condition is TRUE
print("Showing Facebook information")
} else {
  print("Unknown medium")
}

# Control structure for num_views
if (num_views > 15) {
  print("You're popular!")
} else if (num_views <= 15 & num_views > 10) {
  # Add code to print correct string when condition is TRUE
print("Your number of views is average")
} else {
  print("Try to be more visible!")
}
```


## 1-15 Take control!

In this exercise, you will combine everything that you've learned so far: relational operators, logical operators and control constructs. You'll need it all!

We've pre-defined two values for you: li and fb, denoting the number of profile views your LinkedIn and Facebook profile had on the last day of recordings. Go through the instructions to create R code that generates a 'social media score', sms, based on the values of li and fb.

```{r}
# Variables related to your last day of recordings
li <- 15
fb <- 9

# Code the control-flow construct
if (li >= 15 & fb >= 15) {
  sms <- 2 * (li + fb)
} else if (li < 10 & fb < 10) {
  sms <- 0.5 * (li + fb)
} else {
  sms <- li + fb
}

# Print the resulting sms to the console
sms
```



## 2- Loops

Loops can come in handy on numerous occasions. While loops are like repeated if statements, the for loop is designed to iterate over all elements in a sequence. Learn about them in this chapter.

## 2-2 Write a while loop

Let's get you started with building a while loop from the ground up. Have another look at its recipe:

while (condition) {
  expr
}

Remember that the condition part of this recipe should become FALSE at some point during the execution. Otherwise, the while loop will go on indefinitely.

If your session expires when you run your code, check the body of your while loop carefully.

Have a look at the sample code provided; it initializes the speed variables and already provides a while loop template to get you started.

```{r}
# Initialize the speed variable
speed <- 64

# Code the while loop
while ( speed> 30 ) {
  print ("Slow down!")
  speed <- speed-7
}

# Print out the speed variable
speed
```

## 2-3 Throw in more conditionals

In the previous exercise, you simulated the interaction between a driver and a driver's assistant: When the speed was too high, "Slow down!" got printed out to the console, resulting in a decrease of your speed by 7 units.

There are several ways in which you could make your driver's assistant more advanced. For example, the assistant could give you different messages based on your speed or provide you with a current speed at a given moment.

A while loop similar to the one you've coded in the previous exercise is already available for you to use. It prints out your current speed, but there's no code that decreases the speed variable yet, which is pretty dangerous. Can you make the appropriate changes?

```{r}
# Initialize the speed variable
speed <- 64

# Extend/adapt the while loop
while (speed > 30) {
  print(paste("Your speed is",speed))
  if (speed > 48 ) {
    print("Slow down big time!")
    speed <- speed-11
  } else {
    print("Slow down!")
    speed <- speed-6
  }
}
```

## 2-4 Stop the while loop: break

There are some very rare situations in which severe speeding is necessary: what if a hurricane is approaching and you have to get away as quickly as possible? You don't want the driver's assistant sending you speeding notifications in that scenario, right?

This seems like a great opportunity to include the break statement in the while loop you've been working on. Remember that the break statement is a control statement. When R encounters it, the while loop is abandoned completely.

```{r}
# Initialize the speed variable
speed <- 88

while (speed > 30) {
  print(paste("Your speed is", speed))
  
  # Break the while loop when speed exceeds 80
  if (speed > 80) {
    break
  }
  
  if (speed > 48) {
    print("Slow down big time!")
    speed <- speed - 11
  } else {
    print("Slow down!")
    speed <- speed - 6
  }
}
```

## 2-5 Build a while loop from scratch

The previous exercises guided you through developing a pretty advanced while loop, containing a break statement and different messages and updates as determined by control flow constructs. If you manage to solve this comprehensive exercise using a while loop, you're totally ready for the next topic: the for loop.

```{r}
# Initialize i as 1 
i <- 1

# Code the while loop
while (i <= 10) {
  print(3 * i)
  if ((i * 3) %% 8 == 0) {
    break
  }
  i <- i + 1
}
```

## 2-7 Loop over a vector

In the previous video, Filip told you about two different strategies for using the for loop. To refresh your memory, consider the following loops that are equivalent in R:

primes <- c(2, 3, 5, 7, 11, 13)

# loop version 1
for (p in primes) {
  print(p)
}

# loop version 2
for (i in 1:length(primes)) {
  print(primes[i])
}

Remember our linkedin vector? It's a vector that contains the number of views your LinkedIn profile had in the last seven days. The linkedin vector has been pre-defined so that you can fully focus on the instructions!

```{r}
# The linkedin vector has already been defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

# Loop version 1
for (i in linkedin) {
    print(i)
}
print("----------------------")
# Loop version 2
for (i in 1:length(linkedin)) {
    print(linkedin[i])
}
```

## 2-8 Loop over a list

Looping over a list is just as easy and convenient as looping over a vector. There are again two different approaches here:

primes_list <- list(2, 3, 5, 7, 11, 13)

# loop version 1
for (p in primes_list) {
  print(p)
}

# loop version 2
for (i in 1:length(primes_list)) {
  print(primes_list[[i]])
}

Notice that you need double square brackets - [[ ]] - to select the list elements in loop version 2.

Suppose you have a list of all sorts of information on New York City: its population size, the names of the boroughs, and whether it is the capital of the United States. We've already defined a list nyc containing this information (source: Wikipedia).

```{r}
# The nyc list is already specified
nyc <- list(pop = 8405837, 
            boroughs = c("Manhattan", "Bronx", "Brooklyn", "Queens", "Staten Island"), 
            capital = FALSE)

# Loop version 1
for (p in nyc) {
  print(p)
}



# Loop version 2
for (i in 1:length(nyc)) {
  print(nyc[[i]])
}
```

## 2-9 Loop over a matrix

In your workspace, there's a matrix ttt, that represents the status of a tic-tac-toe game. It contains the values "X", "O" and "NA". Print out ttt to get a closer look. On row 1 and column 1, there's "O", while on row 3 and column 2 there's "NA".

To solve this exercise, you'll need a for loop inside a for loop, often called a nested loop. Doing this in R is a breeze! Simply use the following recipe:

for (var1 in seq1) {
  for (var2 in seq2) {
    expr
  }
}

Finish the nested for loops to go over the elements in ttt:

The outer loop should loop over the rows, with loop index i (use 1:nrow(ttt)).

The inner loop should loop over the columns, with loop index j (use 1:ncol(ttt)).

Inside the inner loop, make use of print() and paste() to print out information in the following format: "On row i and column j the board contains x", where x is the value on that position.

```{r}
# The tic-tac-toe matrix ttt has already been defined for you

# define the double for loop
for (i in 1:nrow(ttt)) {
  for (j in 1:ncol(ttt)) {
    print(paste("On row", i, "and column", j, "the board contains",ttt[i,j]))
  }
}
```

## 2-10 Mix it up with control flow

Let's return to the LinkedIn profile views data, stored in a vector linkedin. In the first exercise on for loops you already did a simple printout of each element in this vector. A little more in-depth interpretation of this data wouldn't hurt, right? Time to throw in some conditionals! As with the while loop, you can use the if and else statements inside the for loop.

```{r}
# The linkedin vector has already been defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

# Code the for loop with conditionals
for (li in linkedin) {
  if (li>10 ) {
    print("You're popular!")
  } else {
   print("Be more visible!") 
  }
  print(li)
}
```

## 2-11 Next, you break it

A possible solution to the previous exercise has been provided for you. The code loops over the linkedin vector and prints out different messages depending on the values of li.

In this exercise, you will use the break and next statements:

The break statement abandons the active loop: the remaining code in the loop is skipped and the loop is not iterated over anymore.
The next statement skips the remainder of the code in the loop, but continues the iteration.

Extend the for loop with two new, separate if tests as follows:

If the vector element's value exceeds 16, print out "This is ridiculous, I'm outta here!" and have R abandon the for loop (break).
If the value is lower than 5, print out "This is too embarrassing!" and fast-forward to the next iteration (next).

```{r}
# The linkedin vector has already been defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

# Adapt/extend the for loop
for (li in linkedin) {
  if (li > 10) {
    print("You're popular!")
  } else {
    print("Be more visible!")
  }
  
  # Add if statement with break
  if (li > 16) {
  print("This is ridiculous, I'm outta here!")
  break
  }
  # Add if statement with next
  if (li < 5) {
  print("his is too embarrassing!")
  next
  }
  
  print(li)
}
```

## 2-12 Build a for loop from scratch

This exercise will not introduce any new concepts on for loops.

We already went ahead and defined a variable rquote. This variable has been split up into a vector that contains separate letters and has been stored in a vector chars with the strsplit() function.

Can you write code that counts the number of r's that come before the first u in rquote?

```{r}
# Pre-defined variables
rquote <- "r's internals are irrefutably intriguing"
chars <- strsplit(rquote, split = "")[[1]]

# Initialize rcount
rcount <- 0

# Finish the for loop
for (char in chars) {
  if (char == "r") {
  rcount <- rcount+1
}

if (char == "u") {
  break
}
}
# Print out rcount
print(rcount)
```


## 3- Functions

Functions are an extremely important concept in almost every programming language, and R is no different. Learn what functions are and how to use them—then take charge by writing your own functions.

## 3-2 Function documentation

Before even thinking of using an R function, you should clarify which arguments it expects. All the relevant details such as a description, usage, and arguments can be found in the documentation. To consult the documentation on the sample() function, for example, you can use one of following R commands:

help(sample)
?sample

If you execute these commands, you'll be redirected to www.rdocumentation.org.

A quick hack to see the arguments of the sample() function is the args() function. Try it out in the console:

args(sample)
In the next exercises, you'll be learning how to use the mean() function with increasing complexity. The first thing you'll have to do is get acquainted with the mean() function.

```{r}
# Consult the documentation on the mean() function
?mean

# Inspect the arguments of the mean() function
 args(mean)
```

## 3-3 Use a function

The documentation on the mean() function gives us quite some information:

The mean() function computes the arithmetic mean.
The most general method takes multiple arguments: x and ....
The x argument should be a vector containing numeric, logical or time-related information.
Remember that R can match arguments both by position and by name. Can you still remember the difference? You'll find out in this exercise!

Once more, you'll be working with the view counts of your social network profiles for the past 7 days. These are stored in the linkedin and facebook vectors and have already been created for you.

```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Calculate average number of views
avg_li <- mean(linkedin)
avg_fb <- mean(facebook)


# Inspect avg_li and avg_fb
avg_li
avg_fb
```

## 3-4 Use a function (2)

Check the documentation on the mean() function again:

?mean
The Usage section of the documentation includes two versions of the mean() function. The first usage,

mean(x, ...)
is the most general usage of the mean function. The 'Default S3 method', however, is:

mean(x, trim = 0, na.rm = FALSE, ...)
The ... is called the ellipsis. It is a way for R to pass arguments along without the function having to name them explicitly. The ellipsis will be treated in more detail in future courses.

For the remainder of this exercise, just work with the second usage of the mean function. Notice that both trim and na.rm have default values. This makes them optional arguments.

```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Calculate the mean of the sum
avg_sum <- mean(linkedin+facebook)

# Calculate the trimmed mean of the sum
avg_sum_trimmed <- mean(linkedin+facebook, trim = 0.2)

# Inspect both new variables
avg_sum
avg_sum_trimmed
```

## 3-5 Use a function (3)

In the video, Filip guided you through the example of specifying arguments of the sd() function. The sd() function has an optional argument, na.rm that specified whether or not to remove missing values from the input vector before calculating the standard deviation.

If you've had a good look at the documentation, you'll know by now that the mean() function also has this argument, na.rm, and it does the exact same thing. By default, it is set to FALSE, as the Usage of the default S3 method shows:

mean(x, trim = 0, na.rm = FALSE, ...)
Let's see what happens if your vectors linkedin and facebook contain missing values (NA).

```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, NA, 17, 14)
facebook <- c(17, NA, 5, 16, 8, 13, 14)

# Basic average of linkedin
mean(linkedin)

# Advanced average of linkedin
mean(linkedin, na.rm = TRUE)
```

## 3-6 Functions inside functions

You already know that R functions return objects that you can then use somewhere else. This makes it easy to use functions inside functions, as you've seen before:

speed <- 31

print(paste("Your speed is", speed))

Notice that both the print() and paste() functions use the ellipsis - ... - as an argument. Can you figure out how they're used?

```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, NA, 17, 14)
facebook <- c(17, NA, 5, 16, 8, 13, 14)

# Calculate the mean absolute deviation
mean(abs(linkedin -facebook), na.rm=TRUE)
```

## 3-9 Write your own function

Wow, things are getting serious… you're about to write your own function! Before you have a go at it, have a look at the following function template:

my_fun <- function(arg1, arg2) {
  body
}

Notice that this recipe uses the assignment operator (<-) just as if you were assigning a vector to a variable for example. This is not a coincidence. Creating a function in R basically is the assignment of a function object to a variable! In the recipe above, you're creating a new R variable my_fun, that becomes available in the workspace as soon as you execute the definition. From then on, you can use the my_fun as a function.

```{r}
# Create a function pow_two()
pow_two <- function(a){
a*a
}


# Use the function
pow_two(12)

# Create a function sum_abs()
sum_abs <- function(b,c){
abs(b)+abs(c)
}


# Use the function
sum_abs(-2,3)
```

## 3-10 Write your own function (2)

There are situations in which your function does not require an input. Let's say you want to write a function that gives us the random outcome of throwing a fair die:

throw_die <- function() {
  number <- sample(1:6, size = 1)
  number
}

throw_die()

Up to you to code a function that doesn't take any arguments!

```{r}
# Define the function hello()
hello <- function(){
print("Hi there!")
return(TRUE)
}




# Call the function hello()
hello()
```

## 3-11 Write your own function (3)

Do you still remember the difference between an argument with and without default values? The usage section in the sd() documentation shows the following information:

sd(x, na.rm = FALSE)
This tells us that x has to be defined for the sd() function to be called correctly, however, na.rm already has a default value. Not specifying this argument won't cause an error.

You can define default argument values in your own R functions as well. You can use the following recipe to do so:

my_fun <- function(arg1, arg2 = val2) {
  body
}

The editor on the right already includes an extended version of the pow_two() function from before. Can you finish it?

```{r}
# Finish the pow_two() function
pow_two <- function(x, print_info=TRUE) {
  y <- x ^ 2
  if(print_info == TRUE){
     print(paste(x, "to the power two equals", y))
  }
 
  return(y)
}
```

## 3-14 R you functional?

Now that you've acquired some skills in defining functions with different types of arguments and return values, you should try to create more advanced functions. As you've noticed in the previous exercises, it's perfectly possible to add control-flow constructs, loops and even other functions to your function body.

Remember our social media example? The vectors linkedin and facebook are already defined in the workspace so you can get your hands dirty straight away. As a first step, you will be writing a function that can interpret a single value of this vector. In the next exercise, you will write another function that can handle an entire vector at once.

```{r}

facebook2 <- c(17,  7,  5, 16,  8, 13, 14)
# The linkedin and facebook vectors have already been created for you

# Define the interpret function
interpret <- function(num_views) {
  if (num_views > 15) {
    print("You're popular!")
    return(num_views)
  } else {
    print("Try to be more visible!")
    return(0)
  }
}

# Call the interpret function twice
interpret(linkedin[1])
interpret(facebook2[2])
```

## 3-15 R you functional? (2)

A possible implementation of the interpret() function has been provided for you. In this exercise you'll be writing another function that will use the interpret() function to interpret all the data from your daily profile views inside a vector. Furthermore, your function will return the sum of views on popular days, if asked for. A for loop is ideal for iterating over all the vector elements. The ability to return the sum of views on popular days is something you can code through a function argument with a default value.

```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# The interpret() can be used inside interpret_all()
interpret <- function(num_views) {
  if (num_views > 15) {
    print("You're popular!")
    return(num_views)
  } else {
    print("Try to be more visible!")
    return(0)
  }
}

# Define the interpret_all() function
# views: vector with data to interpret
# return_sum: return total number of views on popular days?
interpret_all <- function(views, return_sum=TRUE) {
  count <- 0

  for (v in views) {
count <- count+ interpret(v)
  }

  if (return_sum==TRUE) {
    count
  } else {
    NULL
  }
}

# Call the interpret_all() function on both linkedin and facebook
interpret_all(linkedin)
interpret_all(facebook)
```

## 3-17 Load an R Package

There are basically two extremely important functions when it comes down to R packages:

install.packages(), which as you can expect, installs a given package.
library() which loads packages, i.e. attaches them to the search list on your R workspace.
To install packages, you need administrator privileges. This means that install.packages() will thus not work in the DataCamp interface. However, almost all CRAN packages are installed on our servers. You can load them with library().

In this exercise, you'll be learning how to load the ggplot2 package, a powerful package for data visualization. You'll use it to create a plot of two variables of the mtcars data frame. The data has already been prepared for you in the workspace.

Before starting, execute the following commands in the console:

search(), to look at the currently attached packages and
qplot(mtcars$wt, mtcars$hp), to build a plot of two variables of the mtcars data frame.
An error should occur, because you haven't loaded the ggplot2 package yet!

```{r}
# Load the ggplot2 package
library(ggplot2)

# Retry the qplot() function
qplot(mtcars$wt, mtcars$hp)

# Check out the currently attached packages again
search()
```


## 4- The apply family

Whenever you're using a for loop, you may want to revise your code to see whether you can use the lapply function instead. Learn all about this intuitive way of applying a function over a list or a vector, and how to use its variants, sapply and vapply.

## 4-2 Use lapply with a built-in R function

Before you go about solving the exercises below, have a look at the documentation of the lapply() function. The Usage section shows the following expression:

lapply(X, FUN, ...)

To put it generally, lapply takes a vector or list X, and applies the function FUN to each of its members. If FUN requires additional arguments, you pass them after you've specified X and FUN (...). The output of lapply() is a list, the same length as X, where each element is the result of applying FUN on the corresponding element of X.

Now that you are truly brushing up on your data science skills, let's revisit some of the most relevant figures in data science history. We've compiled a vector of famous mathematicians/statisticians and the year they were born. Up to you to extract some information!

```{r}
# The vector pioneers has already been created for you
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")

# Split names from birth year
split_math <- strsplit(pioneers, split = ":")

# Convert to lowercase strings: split_low
split_low <- lapply(split_math, tolower)

# Take a look at the structure of split_low
str(split_low)
```

## 4-3 Use lapply with your own function

As Filip explained in the instructional video, you can use lapply() on your own functions as well. You just need to code a new function and make sure it is available in the workspace. After that, you can use the function inside lapply() just as you did with base R functions.

In the previous exercise you already used lapply() once to convert the information about your favorite pioneering statisticians to a list of vectors composed of two character strings. Let's write some code to select the names and the birth years separately.

The sample code already includes code that defined select_first(), that takes a vector as input and returns the first element of this vector.

```{r}
# Code from previous exercise:
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)

# Write function select_first()
select_first <- function(x) {
  x[1]
}

# Apply select_first() over split_low: names
names <- lapply(split_low, select_first)

# Write function select_second()
select_second <- function(x) {
  x[2]
}

# Apply select_second() over split_low: years
years <- lapply(split_low, select_second)
```


## 4-4 lapply and anonymous functions

Writing your own functions and then using them inside lapply() is quite an accomplishment! But defining functions to use them only once is kind of overkill, isn't it? That's why you can use so-called anonymous functions in R.

Previously, you learned that functions in R are objects in their own right. This means that they aren't automatically bound to a name. When you create a function, you can use the assignment operator to give the function a name. It's perfectly possible, however, to not give the function a name. This is called an anonymous function:

# Named function
triple <- function(x) { 3 * x }

# Anonymous function with same implementation
function(x) { 3 * x }

# Use anonymous function inside lapply()
lapply(list(1,2,3), function(x) { 3 * x })

split_low is defined for you.

```{r}
# split_low has been created for you
split_low

# Transform: use anonymous function inside lapply
function(x) {
  x[1]
}
names <- lapply(split_low, function(x) {x[1]})

# Transform: use anonymous function inside lapply
function(x) {
  x[2]
}
years <- lapply(split_low, function(x) {x[2]})
```


## 4-5 Use lapply with additional arguments

In the video, the triple() function was transformed to the multiply() function to allow for a more generic approach. lapply() provides a way to handle functions that require more than one argument, such as the multiply() function:

multiply <- function(x, factor) {
  x * factor
}

lapply(list(1,2,3), multiply, factor = 3)
On the right we've included a generic version of the select functions that you've coded earlier: select_el(). It takes a vector as its first argument, and an index as its second argument. It returns the vector's element at the specified index.

```{r}
# Definition of split_low
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)

# Generic select function
select_el <- function(x, index) {
  x[index]
}

# Use lapply() twice on split_low: names and years
names <- lapply(split_low, select_el, index=1)
years <- lapply(split_low, select_el, index=2)
```


## 4-8 How to use sapply

You can use sapply() similar to how you used lapply(). The first argument of sapply() is the list or vector X over which you want to apply a function, FUN. Potential additional arguments to this function are specified afterwards (...):

sapply(X, FUN, ...)

In the next couple of exercises, you'll be working with the variable temp, that contains temperature measurements for 7 days. temp is a list of length 7, where each element is a vector of length 5, representing 5 measurements on a given day. This variable has already been defined in the workspace: type str(temp) to see its structure.

```{r}
# temp has already been defined in the workspace
temp <- list(c(3, 7, 9, 6, -1),
             c(6, 9, 12, 13, 5),
             c(4, 8, 3, -1, -3),
             c(1, 4, 7, 2, -2),
             c(5, 7, 9, 4, 2),
             c(-3, 5, 8, 9, 4),
             c(3, 6, 9, 4, 1))

# Use lapply() to find each day's minimum temperature
lapply(temp, min)

# Use sapply() to find each day's minimum temperature
sapply(temp, min)

# Use lapply() to find each day's maximum temperature
lapply(temp, max)

# Use sapply() to find each day's maximum temperature
sapply(temp, max)
```


## 4-9 sapply with your own function

Like lapply(), sapply() allows you to use self-defined functions and apply them over a vector or a list:

sapply(X, FUN, ...)

Here, FUN can be one of R's built-in functions, but it can also be a function you wrote. This self-written function can be defined before hand, or can be inserted directly as an anonymous function.

```{r}
# temp is already defined in the workspace

# Finish function definition of extremes_avg
extremes_avg <- function(x) {
  ( min(x) + max(x) ) / 2
}

# Apply extremes_avg() over temp using sapply()
sapply(temp, extremes_avg)

# Apply extremes_avg() over temp using lapply()
lapply(temp, extremes_avg)
```


## 4-10 sapply with function returning vector

In the previous exercises, you've seen how sapply() simplifies the list that lapply() would return by turning it into a vector. But what if the function you're applying over a list or a vector returns a vector of length greater than 1? If you don't remember from the video, don't waste more time in the valley of ignorance and head over to the instructions!

```{r}
# temp is already available in the workspace

# Create a function that returns min and max of a vector: extremes
extremes <- function(x) {
  c(min = min(x), max = max(x))
}

# Apply extremes() over temp with sapply()
sapply(temp, extremes)

# Apply extremes() over temp with lapply()
lapply(temp, extremes)
```


## 4-11 sapply can't simplify, now what?

It seems like we've hit the jackpot with sapply(). On all of the examples so far, sapply() was able to nicely simplify the rather bulky output of lapply(). But, as with life, there are things you can't simplify. How does sapply() react?

We already created a function, below_zero(), that takes a vector of numerical values and returns a vector that only contains the values that are strictly below zero.

```{r}
# temp is already prepared for you in the workspace

# Definition of below_zero()
below_zero <- function(x) {
  return(x[x < 0])
}

# Apply below_zero over temp using sapply(): freezing_s
freezing_s <- sapply(temp, below_zero)

# Apply below_zero over temp using lapply(): freezing_l
freezing_l <- lapply(temp, below_zero)

# Are freezing_s and freezing_l identical?
identical(freezing_s, freezing_l)
```


## 4-12 sapply with functions that return NULL

You already have some apply tricks under your sleeve, but you're surely hungry for some more, aren't you? In this exercise, you'll see how sapply() reacts when it is used to apply a function that returns NULL over a vector or a list.

A function print_info(), that takes a vector and prints the average of this vector, has already been created for you. It uses the cat() function.

```{r}
# temp is already available in the workspace

# Definition of print_info()
print_info <- function(x) {
  cat("The average temperature is", mean(x), "\n")
}

# Apply print_info() over temp using sapply()
sapply(temp, print_info)

# Apply print_info() over temp using lapply()
lapply(temp, print_info)
```


## 4-15 Use vapply

Before you get your hands dirty with the third and last apply function that you'll learn about in this intermediate R course, let's take a look at its syntax. The function is called vapply(), and it has the following syntax:

vapply(X, FUN, FUN.VALUE, ..., USE.NAMES = TRUE)

Over the elements inside X, the function FUN is applied. The FUN.VALUE argument expects a template for the return argument of this function FUN. USE.NAMES is TRUE by default; in this case vapply() tries to generate a named array, if possible.

For the next set of exercises, you'll be working on the temp list again, that contains 7 numerical vectors of length 5. We also coded a function basics() that takes a vector, and returns a named vector of length 3, containing the minimum, mean and maximum value of the vector respectively.

```{r}
# temp is already available in the workspace

# Definition of basics()
basics <- function(x) {
  c(min = min(x), mean = mean(x), max = max(x))
}

# Apply basics() over temp using vapply()
vapply(temp, basics, numeric(3))
```


## 4-16 Use vapply (2)

So far you've seen that vapply() mimics the behavior of sapply() if everything goes according to plan. But what if it doesn't?

In the video, Filip showed you that there are cases where the structure of the output of the function you want to apply, FUN, does not correspond to the template you specify in FUN.VALUE. In that case, vapply() will throw an error that informs you about the misalignment between expected and actual output.

```{r}
# temp is already available in the workspace

# Definition of the basics() function
basics <- function(x) {
  c(min = min(x), mean = mean(x), median = median(x), max = max(x))
}

# Fix the error:
vapply(temp, basics, numeric(4))
```


## 4-17 From sapply to vapply

As highlighted before, vapply() can be considered a more robust version of sapply(), because you explicitly restrict the output of the function you want to apply. Converting your sapply() expressions in your own R scripts to vapply() expressions is therefore a good practice (and also a breeze!).

```{r}
# temp is already defined in the workspace

# Convert to vapply() expression
vapply(temp, max, numeric(1))

# Convert to vapply() expression
vapply(temp, function(x, y) { mean(x) > y }, logical(1), y = 5)
```



## 5- Utilities

Mastering R programming is not only about understanding its programming concepts. Having a solid understanding of a wide range of R functions is also important. This chapter introduces you to many useful functions for data structure manipulation, regular expressions, and working with times and dates.

## 5-2 Mathematical utilities

Have another look at some useful math functions that R features:

abs(): Calculate the absolute value.
sum(): Calculate the sum of all the values in a data structure.
mean(): Calculate the arithmetic mean.
round(): Round the values to 0 decimal places by default. Try out ?round in the console for variations of round() and ways to change the number of digits to round to.

As a data scientist in training, you've estimated a regression model on the sales data for the past six months. After evaluating your model, you see that the training error of your model is quite regular, showing both positive and negative values. A vector errors containing the error values has been pre-defined for you.

```{r}
# The errors vector has already been defined for you
errors <- c(1.9, -2.6, 4.0, -9.5, -3.4, 7.3)

# Sum of absolute rounded values of errors
sum(abs(round(errors)))
```

## 5-3 Find the error

We went ahead and pre-loaded some code for you, but there's still an error. Can you trace it and fix it?

In times of despair, help with functions such as sum() and rev() are a single command away; simply execute the code ?sum and ?rev.

```{r}
# Don't edit these two lines
vec1 <- c(1.5, 2.5, 8.4, 3.7, 6.3)
vec2 <- rev(vec1)
#vec1
#vec2
# Fix the error
#mean(abs(vec1), abs(vec2))
mean(c(abs(vec1), abs(vec2)))
```

## 5-4 Data Utilities

R features a bunch of functions to juggle around with data structures::

seq(): Generate sequences, by specifying the from, to, and by arguments.

rep(): Replicate elements of vectors and lists.

sort(): Sort a vector in ascending order. Works on numerics, but also on character strings and logicals.

rev(): Reverse the elements in a data structures for which reversal is defined.

str(): Display the structure of any R object.
append(): Merge vectors or lists.

is.*(): Check for the class of an R object.

as.*(): Convert an R object from one class to another.

unlist(): Flatten (possibly embedded) lists to produce a
vector.

Remember the social media profile views data? Your LinkedIn and Facebook view counts for the last seven days have been pre-defined as lists.

```{r}
# The linkedin and facebook lists have already been created for you
linkedin <- list(16, 9, 13, 5, 2, 17, 14)
facebook <- list(17, 7, 5, 16, 8, 13, 14)

# Convert linkedin and facebook to a vector: li_vec and fb_vec
li_vec <- as.vector(linkedin)
fb_vec <- as.vector(facebook)

# Append fb_vec to li_vec: social_vec
social_vec <- append(li_vec, fb_vec)

# Sort social_vec
sort(as.numeric(social_vec), decreasing=TRUE)
```

## 5-5 Find the error (2)

Just as before, let's switch roles. It's up to you to see what unforgivable mistakes we've made. Go fix them!

```{r}
# Fix me
#seq(rep(1, 7, by = 2), times = 7)
rep(seq(1, 7, by = 2), times = 7)
```

## 5-6 Beat Gauss using R

There is a popular story about young Gauss. As a pupil, he had a lazy teacher who wanted to keep the classroom busy by having them add up the numbers 1 to 100. Gauss came up with an answer almost instantaneously, 5050. On the spot, he had developed a formula for calculating the sum of an arithmetic series. There are more general formulas for calculating the sum of an arithmetic series with different starting values and increments. Instead of deriving such a formula, why not use R to calculate the sum of a sequence?

```{r}
# Create first sequence: seq1
seq1 <- seq(1,500,by=3)

# Create second sequence: seq2
seq2 <-  seq(1200,900,by=-7)

# Calculate total sum of the sequences
sum(seq1, seq2)
```

## 5-8 grepl & grep

In their most basic form, regular expressions can be used to see whether a pattern exists inside a character string or a vector of character strings. For this purpose, you can use:

grepl(), which returns TRUE when a pattern is found in the corresponding character string.
grep(), which returns a vector of indices of the character strings that contains the pattern.
Both functions need a pattern and an x argument, where pattern is the regular expression you want to match for, and the x argument is the character vector from which matches should be sought.

In this and the following exercises, you'll be querying and manipulating a character vector of email addresses! The vector emails has been pre-defined so you can begin with the instructions straight away!

Use grepl() to generate a vector of logicals that indicates whether these email addressess contain "edu". Print the result to the output.

Do the same thing with grep(), but this time save the resulting indexes in a variable hits.

Use the variable hits to select from the emails vector only the emails that contain "edu".




```{r}
# The emails vector has already been defined for you
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "dalai.lama@peace.org",
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")

# Use grepl() to match for .edu addresses more robustly
grepl(pattern = "@.*\\.edu$", x = emails)

# Use grep() to match for .edu addresses more robustly, save result to hits
hits <- grep(pattern = "@.*\\.edu$", x = emails)

# Subset emails using hits
emails[hits]
```

## 5-9 grepl & grep (2)

You can use the caret, ^, and the dollar sign, $ to match the content located in the start and end of a string, respectively. This could take us one step closer to a correct pattern for matching only the ".edu" email addresses from our list of emails. But there's more that can be added to make the pattern more robust:

@, because a valid email must contain an at-sign.

.*, which matches any character (.) zero or more times (*).

Both the dot and the asterisk are metacharacters. You can use them to match any character between the at-sign and the ".edu" portion of an email address.

\\.edu$, to match the ".edu" part of the email at the end of the string. The \\ part escapes the dot: it tells R that you want to use the . as an actual character.

```{r}
# The emails vector has already been defined for you
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "dalai.lama@peace.org",
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")

# Use grepl() to match for .edu addresses more robustly
grepl(pattern = "@.*\\.edu$", x = emails)

# Use grep() to match for .edu addresses more robustly, save result to hits
hits <- grep(pattern = "@.*\\.edu$", x = emails)

# Subset emails using hits
emails[hits]
```

## 5-10 sub & gsub

While grep() and grepl() were used to simply check whether a regular expression could be matched with a character vector, sub() and gsub() take it one step further: you can specify a replacement argument. If inside the character vector x, the regular expression pattern is found, the matching element(s) will be replaced with replacement.sub() only replaces the first match, whereas gsub() replaces all matches.

Suppose that emails vector you've been working with is an excerpt of DataCamp's email database. Why not offer the owners of the .edu email addresses a new email address on the datacamp.edu domain? This could be quite a powerful marketing stunt: Online education is taking over traditional learning institutions! Convert your email and be a part of the new generation!

With the advanced regular expression "@.*\\.edu$", use sub() to replace the match with "@datacamp.edu". Since there will only be one match per character string, gsub() is not necessary here. Inspect the resulting output.

```{r}
# The emails vector has already been defined for you
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "global@peace.org",
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")

# Use sub() to convert the email domains to datacamp.edu
sub(pattern = "@.*\\.edu$", replacement = "@datacamp.edu", x = emails)
```

## 5-13 Right here, right now

In R, dates are represented by Date objects, while times are represented by POSIXct objects. Under the hood, however, these dates and times are simple numerical values. Date objects store the number of days since the 1st of January in 1970. POSIXct objects on the other hand, store the number of seconds since the 1st of January in 1970.

The 1st of January in 1970 is the common origin for representing times and dates in a wide range of programming languages. There is no particular reason for this; it is a simple convention. Of course, it's also possible to create dates and times before 1970; the corresponding numerical values are simply negative in this case.

```{r}
# Get the current date: today
today <- Sys.Date()

# See what today looks like under the hood
unclass(today)

# Get the current time: now
now <- Sys.time()

# See what now looks like under the hood
unclass(now)
```

## 5-14 Create and format dates

To create a Date object from a simple character string in R, you can use the as.Date() function. The character string has to obey a format that can be defined using a set of symbols (the examples correspond to 13 January, 1982):

%Y: 4-digit year (1982)
%y: 2-digit year (82)
%m: 2-digit month (01)
%d: 2-digit day of the month (13)
%A: weekday (Wednesday)
%a: abbreviated weekday (Wed)
%B: month (January)
%b: abbreviated month (Jan)
The following R commands will all create the same Date object for the 13th day in January of 1982:

as.Date("1982-01-13")
as.Date("Jan-13-82", format = "%b-%d-%y")
as.Date("13 January, 1982", format = "%d %B, %Y")
Notice that the first line here did not need a format argument, because by default R matches your character string to the formats "%Y-%m-%d" or "%Y/%m/%d".

In addition to creating dates, you can also convert dates to character strings that use a different date notation. For this, you use the format() function. Try the following lines of code:

today <- Sys.Date()
format(Sys.Date(), format = "%d %B, %Y")
format(Sys.Date(), format = "Today is a %A!")

```{r}
# Definition of character strings representing dates
str1 <- "May 23, '96"
str2 <- "2012-03-15"
str3 <- "30/January/2006"

# Convert the strings to dates: date1, date2, date3
date1 <- as.Date(str1, format = "%b %d, '%y")
date2 <- as.Date(str2, format = "%Y-%m-%d")
date3 <- as.Date(str3, format = "%d/%b/%Y")

# Convert dates to formatted strings
format(date1, "%A")
format(date2, "%d")
format(date3, "%b %Y")
```

## 5-15 Create and format times

Similar to working with dates, you can use as.POSIXct() to convert from a character string to a POSIXct object, and format() to convert from a POSIXct object to a character string. Again, you have a wide variety of symbols:

%H: hours as a decimal number (00-23)
%I: hours as a decimal number (01-12)
%M: minutes as a decimal number
%S: seconds as a decimal number
%T: shorthand notation for the typical format %H:%M:%S
%p: AM/PM indicator
For a full list of conversion symbols, consult the strptime documentation in the console:

?strptime
Again,as.POSIXct() uses a default format to match character strings. In this case, it's %Y-%m-%d %H:%M:%S. In this exercise, abstraction is made of different time zones.

Convert two strings that represent timestamps, str1 and str2, to POSIXct objects called time1 and time2.

Using format(), create a string from time1 containing only the minutes.

From time2, extract the hours and minutes as "hours:minutes AM/PM". Refer to the assignment text above to find the correct conversion symbols!

```{r}
# Definition of character strings representing times
str1 <- "May 23, '96 hours:23 minutes:01 seconds:45"
str2 <- "2012-3-12 14:23:08"

# Convert the strings to POSIXct objects: time1, time2
time1 <- as.POSIXct(str1, format = "%B %d, '%y hours:%H minutes:%M seconds:%S")
time2 <- as.POSIXct(str2, format = "%Y-%m-%d %H:%M:%S")

# Convert times to formatted strings
format(time1, "%M")
format(time2, "%I:%M %p")
```

## 5-16 Calculations with Dates

Both Date and POSIXct R objects are represented by simple numerical values under the hood. This makes calculation with time and date objects very straightforward: R performs the calculations using the underlying numerical values, and then converts the result back to human-readable time information again.

You can increment and decrement Date objects, or do actual calculations with them:

today <- Sys.Date()
today + 1
today - 1

as.Date("2015-03-12") - as.Date("2015-02-27")
To control your eating habits, you decided to write down the dates of the last five days that you ate pizza. In the workspace, these dates are defined as five Date objects, day1 to day5. A vector pizza containing these 5 Date objects has been pre-defined for you.

```{r}
# day1, day2, day3, day4 and day5 are already available in the workspace
day1 <- as.Date("2021-10-09")
day2 <- as.Date("2021-10-11")
day3 <- as.Date("2021-10-16")
day4 <- as.Date("2021-10-22")
day5 <- as.Date("2021-10-27")
# Difference between last and first pizza day
day5-day1

# Create vector pizza
pizza <- c(day1, day2, day3, day4, day5)

# Create differences between consecutive pizza days: day_diff
day_diff <- diff(pizza)

# Average period between two consecutive pizza days
mean(day_diff)
```

## 5-17 Calculations with Times

Calculations using POSIXct objects are completely analogous to those using Date objects. Try to experiment with this code to increase or decrease POSIXct objects:

now <- Sys.time()
now + 3600          # add an hour
now - 3600 * 24     # subtract a day
Adding or subtracting time objects is also straightforward:

birth <- as.POSIXct("1879-03-14 14:37:23")
death <- as.POSIXct("1955-04-18 03:47:12")
einstein <- death - birth
einstein

You're developing a website that requires users to log in and out. You want to know what is the total and average amount of time a particular user spends on your website. This user has logged in 5 times and logged out 5 times as well. These times are gathered in the vectors login and logout, which are already defined in the workspace.

Calculate the difference between the two vectors logout and login, i.e. the time the user was online in each independent session. Store the result in a variable time_online.

Inspect the variable time_online by printing it.

Calculate the total time that the user was online. Print the result.

Calculate the average time the user was online. Print the result.

```{r}
# login and logout are already defined in the workspace
logout <- as.POSIXct(c("2021-10-13 10:56:29 UTC", "2021-10-18 09:14:52 UTC", "2021-10-18 12:35:48 UTC", "2021-10-18 13:17:22 UTC", "2021-10-20 22:08:47 UTC"))

login <- as.POSIXct(c("2021-10-13 10:18:04 UTC", "2021-10-18 09:14:18 UTC", "2021-10-18 12:21:51 UTC", "2021-10-18 12:37:24 UTC", "2021-10-20 21:37:55 UTC"))

class(logout)

# Calculate the difference between login and logout: time_online
time_online <- logout-login

# Inspect the variable time_online
time_online

# Calculate the total time online
sum(time_online)

# Calculate the average time online
mean(time_online)

```

## 5-18 Time is of the essence

The dates when a season begins and ends can vary depending on who you ask. People in Australia will tell you that spring starts on September 1st. The Irish people in the Northern hemisphere will swear that spring starts on February 1st, with the celebration of St. Brigid's Day. Then there's also the difference between astronomical and meteorological seasons: while astronomers are used to equinoxes and solstices, meteorologists divide the year into 4 fixed seasons that are each three months long. (source: www.timeanddate.com)

A vector astro, which contains character strings representing the dates on which the 4 astronomical seasons start, has been defined on your workspace. Similarly, a vector meteo has already been created for you, with the meteorological beginnings of a season.

Use as.Date() to convert the astro vector to a vector containing Date objects. You will need the %d, %b and %Y symbols to specify the format. Store the resulting vector as astro_dates.

Use as.Date() to convert the meteo vector to a vector with Date objects. This time, you will need the %B, %d and %y symbols for the format argument. Store the resulting vector as meteo_dates.

With a combination of max(), abs() and -, calculate the maximum absolute difference between the astronomical and the meteorological beginnings of a season, i.e. astro_dates and meteo_dates. Simply print this maximum difference to the console output.

```{r}
astro <- as.character(list(spring= "20-Mar-2015" , summer= "25-Jun-2015", fall= "23-Sep-2015", winter= "22-Dec-2015"))

meteo <- as.character(list(spring= "March 1, 15" , summer= "June 1, 15", fall= "September 1, 15", winter= "December 1, 15"))

# Convert astro to vector of Date objects: astro_dates
astro_dates <- as.Date(astro, format = "%d-%b-%Y")

# Convert meteo to vector of Date objects: meteo_dates
meteo_dates <- as.Date(meteo, format = "%B %d, %y")

# Calculate the maximum absolute difference between astro_dates and meteo_dates
max(abs(meteo_dates - astro_dates))
```

