---
title: Time Series Analysis in R
author: ''
date: '2021-10-27'
slug: time-series-analysis-in-r
categories: []
tags:
  - Time Series
---

```{r }
 library(tidyverse)
Nile <- ts(read.csv("Nile.csv"), start = 1871 , end = 1970, frequency = 1)
class(Nile)

data_vector <- c(2.0521941073,  4.2928852797,  3.3294132944,  3.5085950670,  0.0009576938, 1.9217186345,  0.7978134128,  0.2999543435, 0.9435687536,  0.5748283388, -0.0034005903,  0.3448649176,  2.2229761136,  0.1763144576, 2.7097622770, 1.2501948965, -0.4007164754,  0.8852732121, -1.5852420266, -2.2829278891, -2.5609531290, -3.1259963754, -2.8660295895, -1.7847009207, -1.8894912908, -2.7255351194, -2.1033141800, -0.0174256893, -0.3613204151, -2.9008403327, -3.2847440927, -2.8684594718, -1.9505074437, -4.8801892525, -3.2634605353, -1.6396062522, -3.3012575840, -2.6331245433, -1.7058354022, -2.2119825061, -0.5170595186,  0.0752508095, -0.8406994716, -1.4022683487, -0.1382114230, -1.4065954703, -2.3046941055,  1.5073891432, 0.7118679477, -1.1300519022)

eu_stocks <- as.ts(read.csv("eu_stocks.csv"), start = 1998, end = 1998)
#eu_stocks <- ts(eu_stocks1, start = 1998, end = 1998)

rapid_growth <- as.vector(read.csv("rapid_growth.csv"))
z <- read.csv("z.csv")$x
x <- read.csv("x.csv")$x
y <- read.csv("y.csv")$x
```

## Course Description

Many phenomena in our day-to-day lives, such as the movement of stock prices, are measured in intervals over a period of time. Time series analysis methods are extremely useful for analyzing these special data types. In this course, you will be introduced to some core time series analysis concepts and techniques.

## 1- Exploratory time series data analysis

This chapter will give you insights on how to organize and visualize time series data in R. You will learn several simplifying assumptions that are widely used in time series analysis, and common characteristics of financial time series.


## 1-2 Exploring raw time series

The most common first step when conducting time series analysis is to display your time series dataset in a visually intuitive format. The most useful way to view raw time series data in R is to use the print() command, which displays the Start, End, and Frequency of your data along with the observations.

Another useful command for viewing time series data in R is the length() function, which tells you the total number of observations in your data.

Some datasets are very long, and previewing a subset of data is more suitable than displaying the entire series. The head(___, n =___) and tail(___, n =___) functions, in which n is the number of items to display, focus on the first and last few elements of a given dataset respectively.

In this exercise, you'll explore the famous River Nile annual streamflow data, Nile. This time series dataset includes some metadata information. When calling print(Nile), note that Start = 1871 indicates that 1871 is the year of the first annual observation, and End = 1970 indicates 1970 is the year of the last annual observation.

```{r}
# Print the Nile dataset
print(Nile)

# List the number of observations in the Nile dataset
length(Nile)

# Display the first 10 elements of the Nile dataset
head(Nile, n=10)

# Display the last 12 elements of the Nile dataset
tail(Nile, n=12)
```

## 1-3 Basic time series plots

While simple commands such as print(), length(), head(), and tail() provide crucial information about your time series data, another very useful way to explore any data is to generate a plot.

In this exercise, you will plot the River Nile annual streamflow data using the plot() function. For time series data objects such as Nile, a Time index for the horizontal axis is typically included. From the previous exercise, you know that this data spans from 1871 to 1970, and horizontal tick marks are labeled as such. The default label of "Time" is not very informative. Since these data are annual measurements, you should use the label "Year". While you're at it, you should change the vertical axis label to "River Volume (1e9 m^{3})".

Additionally, it helps to have an informative title, which can be set using the argument main. For your purposes, a useful title for this figure would be "Annual River Nile Volume at Aswan, 1871-1970".

Finally, the default plotting type for time series objects is "l" for line. Connecting consecutive observations can help make a time series plot more interpretable. Sometimes it is also useful to include both the observations points as well as the lines, and we instead use "b" for both.

```{r}
# Plot the Nile data
plot(Nile )

# Plot the Nile data with xlab and ylab arguments
plot(Nile , xlab = "Year", ylab = "River Volume (1e9 m^{3})")

# Plot the Nile data with xlab, ylab, main, and type arguments
plot(Nile , xlab = "Year", ylab = "River Volume (1e9 m^{3})", main = "Annual River Nile Volume at Aswan, 1871-1970", type ="b")

```

## 1-4 What does the time index tell us?

Some data are naturally evenly spaced by time. The time series discrete_data shown in the top figure has 20 observations, with one observation appearing at each of the discrete time indices 1 through 20. Discrete time indexing is appropriate for discrete_data.

The time series continuous_series shown in the bottom figure also has 20 observations, it is following the same periodic pattern as discrete_data, but its observations are not evenly spaced. Its first, second, and last observations were observed at times 1.210322, 1.746137, and 20.180524, respectively. Continuous time indexing is natural for continuous_series, however, the observations are approximately evenly spaced, with about 1 observation observed per time unit. Let's investigate using a discrete time indexing for continuous_series.

```{r}

continuous_time_index <- c(1.210322, 1.746137, 2.889634, 3.591384, 5.462065, 5.510933, 7.074295, 8.264398, 9.373382, 9.541063, 11.161122, 12.378371, 13.390559, 14.066280, 15.093547, 15.864515, 16.857413, 18.091457, 19.365451, 20.180524)

continuous_series <- c(0.56889468  , 0.76630408, 0.99207512, 0.97481741, 0.39912320, 0.37660246, -0.38532033, -0.83635852, -0.99966983, -0.99831019, -0.64622280, -0.09386151, 0.40052909, 0.68160578, 0.95318159,0.99693803, 0.83934194, 0.37003754, -0.25509676, -0.61743983)

# Plot the continuous_series using continuous time indexing
par(mfrow=c(2,1))
plot(continuous_time_index,continuous_series, type = "b")

# Make a discrete time index using 1:20 
discrete_time_index <- 1:20 

# Now plot the continuous_series using discrete time indexing
plot(discrete_time_index,continuous_series, type = "b")
```

## 1-6 Identifying the sampling frequency

In addition to viewing your data and plotting over time, there are several additional operations that can be performed on time series datasets.

The start() and end() functions return the time index of the first and last observations, respectively. The time() function calculates a vector of time indices, with one element for each time index on which the series was observed.

The deltat() function returns the fixed time interval between observations and the frequency() function returns the number of observations per unit time. Finally, the cycle() function returns the position in the cycle of each observation.

In this exercise, you'll practice applying these functions to the AirPassengers dataset, which reports the monthly total international airline passengers (in thousands) from 1949 to 1960.

```{r}
head(AirPassengers,3)
class(AirPassengers)
# Plot AirPassengers
plot(AirPassengers)

# View the start and end dates of AirPassengers
start(AirPassengers)
end(AirPassengers)

# Use time(), deltat(), frequency(), and cycle() with AirPassengers 
time(AirPassengers)
deltat(AirPassengers)
frequency(AirPassengers)
cycle(AirPassengers)

```

## 1-8 Missing values

Sometimes there are missing values in time series data, denoted NA in R, and it is useful to know their locations. It is also important to know how missing values are handled by various R functions. Sometimes we may want to ignore any missingness, but other times we may wish to impute or estimate the missing values.

Let's again consider the monthly AirPassengers dataset, but now the data for the year 1956 are missing. In this exercise, you'll explore the implications of this missing data and impute some new data to solve the problem.

The mean() function calculates the sample mean, but it fails in the presence of any NA values. Use mean(___, na.rm = TRUE) to calculate the mean with all missing values removed. It is common to replace missing values with the mean of the observed values. Does this simple data imputation scheme appear adequate when applied the the AirPassengers dataset?

```{r}
# Plot the AirPassengers data
plot(AirPassengers)

# Compute the mean of AirPassengers
mean(AirPassengers, na.rm=TRUE)

# Impute mean values to NA in AirPassengers
AirPassengers[85:96] <- mean(AirPassengers, na.rm = TRUE)

# Generate another plot of AirPassengers
plot(AirPassengers)

# Add the complete AirPassengers data to your plot
rm(AirPassengers)
points(AirPassengers, type = "l", col = 2, lty = 3)

```


## 1-10 Creating a time series object with ts()

The function ts() can be applied to create time series objects. A time series object is a vector (univariate) or matrix (multivariate) with additional attributes, including time indices for each observation, the sampling frequency and time increment between observations, and the cycle length for periodic data. Such objects are of the ts class, and represent data that has been observed at (approximately) equally spaced time points. Now you will create time series objects yourself.

The advantage of creating and working with time series objects of the ts class is that many methods are available for utilizing time series attributes, such as time index information. For example, as you've seen in earlier exercises, calling plot() on a ts object will automatically generate a plot over time.

In this exercise, you'll familiarize yourself with the ts class by encoding some time series data (saved as data_vector) into ts and exploring the result. Your time series data_vector starts in the year 2004 and has 4 observations per year (i.e. it is quarterly data).

```{r}
# Use print() and plot() to view data_vector
print(data_vector)
plot(data_vector)

# Convert data_vector to a ts object with start = 2004 and frequency = 4
time_series <- ts(data_vector, start = 2004, frequency = 4)

# Use print() and plot() to view time_series
print(time_series)
plot(time_series)
```


## 1-11 Testing whether an object is a time series

When you work to create your own datasets, you can build them as ts objects. Recall the dataset data_vector from the previous exercise, which was just a vector of numbers, and time_series, the ts object you created from data_vector using the ts() function and information regarding the start time and the observation frequency. As a reminder, data_vector and time_series are shown in the plot on the right.

When you use datasets from others, such as those included in an R package, you can check whether they are ts objects using the is.ts() command. The result of the test is either TRUE when the data is of the ts class, or FALSE if it is not.

In this exercise, you'll explore the class of the datasets you've been using throughout this chapter.

```{r}
# Check whether data_vector and time_series are ts objects
is.ts(data_vector)
is.ts(time_series)

# Check whether Nile is a ts object
is.ts(Nile)

# Check whether AirPassengers is a ts object
is.ts(AirPassengers)

```


## 1-12 Plotting a time series object

It is often very useful to plot data we are analyzing, as is the case when conducting time series analysis. If the dataset under study is of the ts class, then the plot() function has methods that automatically incorporate time index information into a figure.

Let's consider the eu_stocks dataset (available in R by default as EuStockMarkets). This dataset contains daily closing prices of major European stock indices from 1991-1998, specifically, from Germany (DAX), Switzerland (SMI), France (CAC), and the UK (FTSE). The data were observed when the markets were open, so there are no observations on weekends and holidays. We will proceed with the approximation that this dataset has evenly spaced observations and is a four dimensional time series.

To conclude this chapter, this exercise asks you to apply several of the functions you've already learned to this new dataset.

```{r}
# Check whether eu_stocks is a ts object
is.ts(eu_stocks)

# View the start, end, and frequency of eu_stocks
start(eu_stocks)
end(eu_stocks)
frequency(eu_stocks)



# Generate a simple plot of eu_stocks
plot(eu_stocks)

# Use ts.plot with eu_stocks
ts.plot(eu_stocks, col = 1:4, xlab = "Year", ylab = "Index Value", main = "Major European Stock Indices, 1991-1998")

# Add a legend to your ts.plot
legend("topleft", colnames(eu_stocks), lty = 1, col = 1:4, bty = "n")

```



## 2- Predicting the future

In this chapter, you will conduct some trend spotting, and learn the white noise (WN) model, the random walk (RW) model, and the definition of stationary processes.

## 2-4 Removing trends in variability via the logarithmic transformation

The logarithmic function log() is a data transformation that can be applied to positively valued time series data. It slightly shrinks observations that are greater than one towards zero, while greatly shrinking very large observations. This property can stabilize variability when a series exhibits increasing variability over time. It may also be used to linearize a rapid growth pattern over time.

The time series rapid_growth has already been loaded, and is shown in the figure on the right. Note the vertical range of the data.

```{r}
 ts.plot(rapid_growth)
# Log rapid_growth
linear_growth <- log(rapid_growth)
  
# Plot linear_growth using ts.plot()
 ts.plot(linear_growth)

```


## 2-5 Removing trends in level by differencing

The first difference transformation of a time series  consists of the differences (changes) between successive observations over time, that is .

Differencing a time series can remove a time trend. The function diff() will calculate the first difference or change series. A difference series lets you examine the increments or changes in a given time series. It always has one fewer observations than the original series.

The time series z has already been loaded, and is shown in the figure on the right.

```{r}
ts.plot(z)
# Generate the first difference of z
dz <- diff(z)
  
# Plot dz
ts.plot(dz)

# View the length of z and dz, respectively
length(z)
length(dz)
```


## 2-6 Removing seasonal trends with seasonal differencing

For time series exhibiting seasonal trends, seasonal differencing can be applied to remove these periodic patterns. For example, monthly data may exhibit a strong twelve month pattern. In such situations, changes in behavior from year to year may be of more interest than changes from month to month, which may largely follow the overall seasonal pattern.

The function diff(..., lag = s) will calculate the lag s difference or length s seasonal change series. For monthly or quarterly data, an appropriate value of s would be 12 or 4, respectively. The diff() function has lag = 1 as its default for first differencing. Similar to before, a seasonally differenced series will have s fewer observations than the original series.

```{r}
# Generate a diff of x with lag = 4. Save this to dx
dx <- diff(x, lag = 4)
  ts.plot(x)
# Plot dx
  ts.plot(dx)

# View the length of x and dx, respectively 
length(x)
length(dx)
```


## 2-8 Simulate the white noise model

The white noise (WN) model is a basic time series model. It is also a basis for the more elaborate models we will consider. We will focus on the simplest form of WN, independent and identically distributed data.

The arima.sim() function can be used to simulate data from a variety of time series models. ARIMA is an abbreviation for the autoregressive integrated moving average class of models we will consider throughout this course.

An ARIMA(p, d, q) model has three parts, the autoregressive order p, the order of integration (or differencing) d, and the moving average order q. We will detail each of these parts soon, but for now we note that the ARIMA(0, 0, 0) model, i.e., with all of these components zero, is simply the WN model.

In this exercise, you will practice simulating a basic WN model.

```{r}
# Simulate a WN model with list(order = c(0, 0, 0))
white_noise <- arima.sim(model = list(order = c(0, 0, 0)), n = 100)

# Plot your white_noise data
ts.plot(white_noise)

# Simulate from the WN model with: mean = 100, sd = 10
white_noise_2 <- arima.sim(model = list(order = c(0, 0, 0)), n = 100, mean = 100, sd = 10)

# Plot your white_noise_2 data
ts.plot(white_noise_2)

```


## 2-9 Estimate the white noise model

For a given time series y we can fit the white noise (WN) model using the arima(..., order = c(0, 0, 0)) function. Recall that the WN model is an ARIMA(0,0,0) model. Applying the arima() function returns information or output about the estimated model. For the WN model this includes the estimated mean, labeled intercept, and the estimated variance, labeled sigma^2.

In this exercise, you'll explore the qualities of the WN model. What is the estimated mean? Compare this with the sample mean using the mean() function. What is the estimated variance? Compare this with the sample variance using the var() function.

The time series y has already been loaded, and is shown in the adjoining figure.

```{r}
# Fit the WN model to y using the arima command
arima(y, order =c(0, 0, 0))

# Calculate the sample mean and sample variance of y
mean(y)
var(y)

```


## 2-11 Simulate the random walk model

The random walk (RW) model is also a basic time series model. It is the cumulative sum (or integration) of a mean zero white noise (WN) series, such that the first difference series of a RW is a WN series. Note for reference that the RW model is an ARIMA(0, 1, 0) model, in which the middle entry of 1 indicates that the model's order of integration is 1.

The arima.sim() function can be used to simulate data from the RW by including the model = list(order = c(0, 1, 0)) argument. We also need to specify a series length n. Finally, you can specify a sd for the series (increments), where the default value is 1.

```{r}
# Generate a RW model using arima.sim
random_walk <- arima.sim(model = list(order = c(0, 1, 0)), n = 100)

# Plot random_walk
ts.plot(random_walk)

# Calculate the first difference series
random_walk_diff <- diff(random_walk)

# Plot random_walk_diff
ts.plot(random_walk_diff)
```


## 2-12 Simulate the random walk model with a drift

A random walk (RW) need not wander about zero, it can have an upward or downward trajectory, i.e., a drift or time trend. This is done by including an intercept in the RW model, which corresponds to the slope of the RW time trend.

For an alternative formulation, you can take the cumulative sum of a constant mean white noise (WN) series, such that the mean corresponds to the slope of the RW time trend.

To simulate data from the RW model with a drift you again use the arima.sim() function with the model = list(order = c(0, 1, 0)) argument. This time, you should add the additional argument mean = ... to specify the drift variable, or the intercept.

```{r}
# Generate a RW model with a drift uing arima.sim
rw_drift <- arima.sim(model = list(order = c(0, 1, 0)), n = 100, mean = 1)

# Plot rw_drift
ts.plot(rw_drift)

# Calculate the first difference series
rw_drift_diff <- diff(rw_drift)

# Plot rw_drift_diff
ts.plot(rw_drift_diff)

```


## 2-13 Estimate the random walk model

For a given time series y we can fit the random walk model with a drift by first differencing the data, then fitting the white noise (WN) model to the differenced data using the arima() command with the order = c(0, 0, 0)) argument.

The arima() command displays information or output about the fitted model. Under the Coefficients: heading is the estimated drift variable, named the intercept. Its approximate standard error (or s.e.) is provided directly below it. The variance of the WN part of the model is also estimated under the label sigma^2.

```{r}
# Difference your random_walk data
rw_diff <- diff(random_walk)

# Plot rw_diff
ts.plot(rw_diff)

# Now fit the WN model to the differenced data
model_wn <- arima(x = rw_diff, order=c(0, 0, 0))

# Store the value of the estimated time trend (intercept)
int_wn <- model_wn$coef

# Plot the original random_walk data
ts.plot(random_walk)

# Use abline(0, ...) to add time trend to the figure
abline(0, int_wn)
```


## 2-16 Are the white noise model or the random walk model stationary?

The white noise (WN) and random walk (RW) models are very closely related. However, only the RW is always non-stationary, both with and without a drift term. This is a simulation exercise to highlight the differences.

Recall that if we start with a mean zero WN process and compute its running or cumulative sum, the result is a RW process. The cumsum() function will make this transformation for you. Similarly, if we create a WN process, but change its mean from zero, and then compute its cumulative sum, the result is a RW process with a drift.

```{r}

```



## 3- Correlation analysis and the autocorrelation function

In this chapter, you will review the correlation coefficient, use it to compare two time series, and also apply it to compare a time series with its past, as an autocorrelation. You will discover the autocorrelation function (ACF) and practice estimating and visualizing autocorrelations for time series data.









## 4- Autoregression

In this chapter, you will learn the autoregressive (AR) model and several of its basic properties. You will also practice simulating and estimating the AR model in R, and compare the AR model with the random walk (RW) model.









## 5- A simple moving average

In this chapter, you will learn the simple moving average (MA) model and several of its basic properties. You will also practice simulating and estimating the MA model in R, and compare the MA model with the autoregressive (AR) model.










```{r}
library(tidyverse)
e <- read.csv("e2.csv")
names(e)
```


```{r}
e %>%  pivot_longer(c(Lab.1...10,Lab.2...10)) %>% 
  ggplot(aes(value,color=name))+geom_density()
```

```{r}
e %>% mutate(q3=Quiz.3..15*20/15) %>% pivot_longer(c(Quiz.1..20,Quiz.2..20,q3)) %>% 
  ggplot(aes(value,color=name))+geom_density()
```

```{r}
Q1_m <- 5*mean(e$Quiz.1..20, na.rm=TRUE)
print(paste("quiz 1: ",Q1_m))

Q2_m <- 5*mean(e$Quiz.2..20, na.rm=TRUE)
print(paste("quiz 2: ",Q2_m))

Q3_m <- 5*20/15*mean(e$Quiz.3..15, na.rm=TRUE)
print(paste("quiz 2: ",Q3_m))
```

```{r}
Q1_max <- min(e$Quiz.1..20, na.rm=TRUE)
print(paste("quiz 1: ",Q1_max))

Q2_max <- min(e$Quiz.2..20, na.rm=TRUE)
print(paste("quiz 2: ",Q2_max))

Q3_max <- 20/15*min(e$Quiz.3..15, na.rm=TRUE)
print(paste("quiz 3: ",Q3_max))
```

