---
title: Cleaning Data in R
author: ''
date: '2021-09-18'
slug: cleaning-data-in-r
categories: []
tags:
  - data_visualization
  - data_wrangling
---

```{r include=FALSE}
library(tidyverse)
library(assertive)

bike_share_rides <- readRDS("bike_share_rides_ch1_1.rds")
ch3_1_accounts <- readRDS("ch3_1_accounts.rds")
fodors <- readRDS("fodors.rds")
sfo_survey <- readRDS("sfo_survey_ch2_1.rds")
zagat <- readRDS("zagat.rds")

dest_sizes <- data.frame(dest_size= c("Small", "Medium", "Large", "Hub"),
                         passengers_per_day= c("0-20K", "20K-70K", "70K-100K", "100K+"))
```

It's commonly said that data scientists spend 80% of their time cleaning and manipulating data and only 20% of their time analyzing it. The time spent cleaning is vital since analyzing dirty data can lead you to draw inaccurate conclusions.

In this course, you'll learn how to clean dirty data. Using R, you'll learn how to identify values that don't look right and fix dirty data by converting data types, filling in missing values, and using fuzzy string matching. As you learn, you’ll brush up on your skills by working with real-world datasets, including bike-share trips, customer asset portfolios, and restaurant reviews—developing the skills you need to go from raw data to awesome insights as quickly and accurately as possible!


## 1-3 Converting data types

Throughout this chapter, you'll be working with San Francisco bike share ride data called bike_share_rides. It contains information on start and end stations of each trip, the trip duration, and some user information.

Before beginning to analyze any dataset, it's important to take a look at the different types of columns you'll be working with, which you can do using glimpse().

In this exercise, you'll take a look at the data types contained in bike_share_rides and see how an incorrect data type can flaw your analysis.

dplyr and assertive are loaded and bike_share_rides is available.

```{r}
# Glimpse at bike_share_rides
glimpse(bike_share_rides)

# Summary of user_birth_year
summary(bike_share_rides$user_birth_year)
```

```{r}
# Convert user_birth_year to factor: user_birth_year_fct
bike_share_rides <- bike_share_rides %>%
  mutate(user_birth_year_fct = factor(user_birth_year))

# Assert user_birth_year_fct is a factor
assert_is_factor(bike_share_rides$user_birth_year_fct)

# Summary of user_birth_year_fct
summary(bike_share_rides$user_birth_year_fct)
```


## 1-4 Trimming strings

In the previous exercise, you were able to identify the correct data type and convert user_birth_year to the correct type, allowing you to extract counts that gave you a bit more insight into the dataset.

Another common dirty data problem is having extra bits like percent signs or periods in numbers, causing them to be read in as characters. In order to be able to crunch these numbers, the extra bits need to be removed and the numbers need to be converted from character to numeric. In this exercise, you'll need to convert the duration column from character to numeric, but before this can happen, the word "minutes" needs to be removed from each value.

dplyr, assertive, and stringr are loaded and bike_share_rides is available.

```{r}
bike_share_rides <- bike_share_rides %>%
  # Remove 'minutes' from duration: duration_trimmed
  mutate(duration_trimmed = str_remove(duration, "minutes"),
         # Convert duration_trimmed to numeric: duration_mins
         duration_mins = as.numeric(duration_trimmed))

# Glimpse at bike_share_rides
glimpse(bike_share_rides)

# Assert duration_mins is numeric
assert_is_numeric(bike_share_rides$duration_mins)

# Calculate mean duration
mean(bike_share_rides$duration_mins)
```

## 1-6 Ride duration constraints

Values that are out of range can throw off an analysis, so it's important to catch them early on. In this exercise, you'll be examining the duration_min column more closely. Bikes are not allowed to be kept out for more than 24 hours, or 1440 minutes at a time, but issues with some of the bikes caused inaccurate recording of the time they were returned.

In this exercise, you'll replace erroneous data with the range limit (1440 minutes), however, you could just as easily replace these values with NAs.

dplyr, assertive, and ggplot2 are loaded and bike_share_rides is available.

```{r}
# Create breaks
breaks <- c(min(bike_share_rides$duration_mins), 0, 1440, max(bike_share_rides$duration_mins))

# Create a histogram of duration_min
ggplot(bike_share_rides, aes(duration_mins)) +
  geom_histogram(breaks = breaks)

# duration_min_const: replace vals of duration_min > 1440 with 1440
bike_share_rides <- bike_share_rides %>%
  mutate(duration_min_const = replace(duration_mins, duration_mins > 1440, 1440))

# Make sure all values of duration_min_const are between 0 and 1440
assert_all_are_in_closed_range(bike_share_rides$duration_min_const, lower = 0, upper = 1440)
```

## 1-7 Back to the future

Something has gone wrong and it looks like you have data with dates from the future, which is way outside of the date range you expected to be working with. To fix this, you'll need to remove any rides from the dataset that have a date in the future. Before you can do this, the date column needs to be converted from a character to a Date. Having these as Date objects will make it much easier to figure out which rides are from the future, since R makes it easy to check if one Date object is before (<) or after (>) another.

dplyr and assertive are loaded and bike_share_rides is available.

```{r warning=FALSE}
library(lubridate)
# Convert date to Date type
bike_share_rides <- bike_share_rides %>%
  mutate(date = as.Date(date))

# Make sure all dates are in the past
assert_all_are_in_past(bike_share_rides$date)

# Filter for rides that occurred before or on today's date
bike_share_rides_past <- bike_share_rides %>%
  filter(date < today())

# Make sure all dates from bike_share_rides_past are in the past
assert_all_are_in_past(bike_share_rides_past$date)
```

## 1-9 Full duplicates

You've been notified that an update has been made to the bike sharing data pipeline to make it more efficient, but that duplicates are more likely to be generated as a result. To make sure that you can continue using the same scripts to run your weekly analyses about ride statistics, you'll need to ensure that any duplicates in the dataset are removed first.

When multiple rows of a data frame share the same values for all columns, they're full duplicates of each other. Removing duplicates like this is important, since having the same value repeated multiple times can alter summary statistics like the mean and median. Each ride, including its ride_id should be unique.

dplyr is loaded and bike_share_rides is available.

```{r}
# Count the number of full duplicates
sum(duplicated(bike_share_rides))

# Remove duplicates
bike_share_rides_unique <- filter(distinct(bike_share_rides))

# Count the full duplicates in bike_share_rides_unique
sum(duplicated(bike_share_rides_unique))
```

## 1-10 Removing partial duplicates

Now that you've identified and removed the full duplicates, it's time to check for partial duplicates. Partial duplicates are a bit tricker to deal with than full duplicates. In this exercise, you'll first identify any partial duplicates and then practice the most common technique to deal with them, which involves dropping all partial duplicates, keeping only the first.

dplyr is loaded and bike_share_rides is available.

```{r}
# Find duplicated ride_ids
bike_share_rides %>% 
  # Count the number of occurrences of each ride_id
  count(ride_id) %>% 
  # Filter for rows with a count > 1
  filter(n > 1)

# Remove full and partial duplicates
bike_share_rides_unique <- bike_share_rides %>%
  # Only based on ride_id instead of all cols
  distinct(ride_id, .keep_all = TRUE)

# Find duplicated ride_ids in bike_share_rides_unique
bike_share_rides_unique %>%
  # Count the number of occurrences of each ride_id
  count(ride_id) %>%
  # Filter for rows with a count > 1
  filter(n > 1)
```

## 1-11 Aggregating partial duplicates

Another way of handling partial duplicates is to compute a summary statistic of the values that differ between partial duplicates, such as mean, median, maximum, or minimum. This can come in handy when you're not sure how your data was collected and want an average, or if based on domain knowledge, you'd rather have too high of an estimate than too low of an estimate (or vice versa).

```{r}
bike_share_rides %>%
  # Group by ride_id and date
  group_by(ride_id, date) %>%
  # Add duration_min_avg column
  mutate(duration_min_avg = mean(duration_mins) ) %>%
  # Remove duplicates based on ride_id and date, keep all cols
  distinct(ride_id, date, .keep_all = TRUE) %>%
  # Remove duration_min column
  select(-duration_mins)
```

## 2-3 Not a member

Now that you've practiced identifying membership constraint problems, it's time to fix these problems in a new dataset. Throughout this chapter, you'll be working with a dataset called sfo_survey, containing survey responses from passengers taking flights from San Francisco International Airport (SFO). Participants were asked questions about the airport's cleanliness, wait times, safety, and their overall satisfaction.

There were a few issues during data collection that resulted in some inconsistencies in the dataset. In this exercise, you'll be working with the dest_size column, which categorizes the size of the destination airport that the passengers were flying to. A data frame called dest_sizes is available that contains all the possible destination sizes. Your mission is to find rows with invalid dest_sizes and remove them from the data frame.

```{r}
# Count the number of occurrences of dest_size
sfo_survey %>%
  count(dest_size)
```

```{r}
# Find bad dest_size rows
sfo_survey %>% 
  # Join with dest_sizes data frame to get bad dest_size rows
   anti_join(dest_sizes, by = "dest_size") %>%
  # Select id, airline, destination, and dest_size cols
  select(id, airline, destination, dest_size)
```


## 2-5 

```{r}

```

## 2-6 

```{r}

```

## 2-7 

```{r}

```

## 2-9 

```{r}

```

## 2-10 

```{r}

```

## 2-11 

```{r}

```

